---
title: "Stat 33B - Lecture Notes 8"
date: October 11, 2020
output: pdf_document
---


For-loops
=========

A for-loop runs a block of code once for each element of a vector or list:
```{r}

```
The idea is the same as for-loops in other languages.

Curly braces `{ }` are only required if you have multiple lines of code.



Use `break` to exit a loop early:
```{r}

```


Use `next` to skip to the next iteration early:
```{r}

```



Loop Indices
============

If you need indices, using `1:n` can cause bugs:
```{r}

```

Use `seq_len(n)` instead:
```{r}

```

Similarly, using `1:length(x)` can cause bugs:
```{r}

```

Use `seq_along(x)` instead:
```{r}

```

More generally, use `seq()` to produce sequences of indices.



While-loops
===========

A while-loop runs a block of code repeatedly while some condition is `TRUE`.

The condition is checked before each iteration:
```{r}

```
Again, the same idea as in other languages.



Some languages have a do-while-loop, which checks the condition after each
iteration (so the first iteration always runs).

R has `repeat`, which is the same as `while (TRUE)`.

You can create a do-while-loop in R with `repeat`:
```{r}

```

These kinds of loops are most useful when the number of iterations is unknown
before the loop runs.



Preallocation
=============

_Preallocation_ means allocating memory for results before a computation.

These functions allocate vectors:

* `character()`
* `complex()`
* `numeric()`
* `logical()`

* `vector()`
* `rep()`

Examples:
```{r}

```

Preallocation is especially important for loops:
```{r}

```


Loops Example
=============

Let's simulate a two-dimensional discrete random walk.

In a random walk, at each time step the walker randomly moves 1 unit along the
x-coordinates or the y-coordinates.

The algorithm is:

1. Flip a coin to determine whether walker moves in x or y.
2. Flip a coin to determine whether walker moves `-1` or `+1`.
3. Update the walker's position.
4. Repeat 1-3 for the desired number of steps.

We can do this with a for-loop:
```{r}

```



Recursion
=========

A _recursive_ function is one that calls itself.

In R, recursion is usually the slowest way to write a loop.

But some problems are naturally recursive, so for those problems it may make
sense to use recursion.

For example:

The Fibonacci numbers are recursive by definition; each Fibonacci number is the
sum of the previous two.

The first few Fibonacci numbers are:
```
1 1 2 3 5 8 13
```

Suppose we want a function that computes the nth Fibonacci number:
```{r}

```
The recursion creates a kind of loop.

This solution for computing the Fibonacci numbers is not efficient.



A better solution is to keep track of the numbers that have already been
computed:
```{r}

```
Using a for-loop would be even more efficient here.


R's built-in `Recall()` function will re-call the enclosing function:
```{r}

```

Without `Recall()`, renaming a recursive function can break the function.



Developing Iterative Code
=========================

See the slides.


When thinking about writing a loop, try (in order):

1. vectorization
2. apply functions
    * Try an apply function if iterations are independent.
3. for/while-loops
    * Try a for-loop if some iterations depend on others.
    * Try a while-loop if the number of iterations is unknown.
4. recursion
    * Convenient for naturally recursive problems (like Fibonacci),
      but often there are faster solutions.


Before you write the loop, try writing the code for just 1 iteration.

Make sure that code works; it's easy to test code for 1 iteration.

When you have 1 iteration working, then try using the code in a loop
(you will have to make some small changes).

If your loop doesn't work, try to figure out which iteration is
causing the problem. One way to do this is to use message() to print
out information.

Then try to write the code for the broken iteration, get that
iteration working, and repeat this whole process.
