---
title: "Stat 33B - Lecture Notes 14"
date: November 29, 2020
output: pdf_document
---


Relational Data
===============

Data split across multiple tables are called _relational data_.

A column shared by several tables is called a _key_.

For example, a grocery store's inventory system might have:

* A table that lists stores
* A table that lists items (fruits, vegetables, etc)
* A table that lists quantity of each item at each store



A _join_ combines two separate tables by based on a common key.

For each row in the first table, the key is compared against the rows
in the second table.

If the key matches, the rows are combined into a new row.

Note that a join can duplicate rows!


Four kinds of joins are:

* Inner: keep matches
* Full: keep everything
* Left: keep all left, keep matching right
* Right: keep matching left, keep all right


Joins are not just used in R; joins are especially important to understand for
working with databases and SQL.









Joins (with dplyr)
==================

We'll use the grocery data set:
```{r}

```


The dplyr package provides several join functions.


The `inner_join()` function only keeps rows where the key matched:
```{r}

```

Use the `by` parameter to specify the key:
```{r}

```


The `left_join()` function keeps all rows in the "left" or first
table.

Rows in the right table are only kept if they match:
```{r}

```


Conversely, the `right_join()` function keeps all rows in the "right"
or second table.

```{r}

```


The `full_join()` function keeps all rows, regardless of whether the key
matched.

As a result, some rows will contain missing values:
```{r}

```


The dplyr package also provides other kinds of joins that are not as common
outside of R. See the docs for details.
