---
title: "Stat 33B - Lecture Notes 11"
date: November 1, 2020
output: pdf_document
---


Attributes
==========

_Attributes_ are name-value pairs in the metadata of an R object.

Some common attributes:

* `class` -- class
* `row.names` -- row names
* `names` -- element names or column names
* `dim` -- dimensions (on matrices)
* `dimnames` -- names of dimensions (on matrices)

There are helper functions like `class` to get and set these:
```{r}

```

There's also a more general way to get and set attributes.

Use the `attributes` function to get all attributes on an object:
```{r}

```

Use the `attr` function to get or set an attribute:
```{r}

```


Use the `structure` function to set all attributes on an object:
```{r}

```

Vectors usually don't have attributes:
```{r}

```

But the `class` function still returns a class:
```{r}

```

The `attr` function shows that the class for this object is not an attribute:
```{r}

```

Most of the time, you should use helper functions to get/set attributes.



The S3 Object System
====================

The `class` function returns an object's S3 classes:
```{r}

```


Use `inherits` to test whether an object is an instance of a specific S3 class:
```{r}

```

A _generic function_ is a special function that has multiple implementations.
These implementations are called _methods_.

When you call a generic function, an appropriate method is selected based on
the class of an argument (usually the first argument). This is called _method
dispatch_, or just _dispatch_.

Use `methods` to list the methods defined for a given generic function:
```{r}

```
This also lists methods for the S4 system.

If you only want S3 methods, use `.S3methods`:
```{r}

```

The pattern for S3 method names is `GENERIC.CLASS`, where:

* `GENERIC` is the name of the generic function.
* `CLASS` is the name of the class that corresponds to the method.

For example: `print.lm` is the `print` method for the `lm` class.

```{r}

```
Methods in packages are not always accessible from the global workspace.

Use `getS3method` or `getAnywhere` to get these methods.

```{r}

```


Not all functions with a `.` in their name are methods.

For example, the `row.names` function is **NOT** a method.

In fact, it's a generic function:
```{r}

```



S3 Objects
==========

An S3 object is just an object with a `class` attribute.

So you can use `class`, `attr`, or `structure` to create an S3 object:
```{r}

```

S3 objects are created from other objects.

Most people create S3 objects from lists.

Then the elements of the list are called _fields_ of the class:
```{r}

```

A _constructor function_ is a function that creates a classed object.

Generally, constructor functions are named after the class of object they
create:
```{r}

```

For an S3 class, it's idiomatic to provide a constructor function.

The constructor function serves as an implicit definition of the class.



S3 Generics & Methods
=====================

Use the `UseMethod` function to create an S3 generic function:
```{r}

```

Use `GENERIC.CLASS` as the name to create a method:
```{r}

```

S3 generics select a method based on one argument.

The default is the first argument. You can optionally specify which argument to
use in `UseMethod`:
```{r}

```


Use `NextMethod` to call the same method for an object's parent class:
```{r}

```



Other Object Systems
====================

See the slides.
