---
title: "STAT 33B Workbook 10"
date: "Nov 5, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This workbook is due __Nov 5, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

If you have any last-minute trouble knitting, **DON'T PANIC**. Submit your Rmd
file on time and follow up in office hours or on Piazza to sort out the PDF.


Attributes
==========

Watch the "Attributes" lecture video.

No exercises for this section.



The S3 Object System
====================

Watch the "The S3 Object System" lecture video.

No exercises for this section.



S3 Objects
==========

Watch the "S3 Objects" lecture video.

The goal of the next 2 exercises is to create an S3 class `polynomial` that can
represent polynomials of any degree.


## Exercise 1

Create a function `poly` that constructs an S3 object with class `polynomial`.
The `poly` function should accepts any number of coefficients as arguments, in
order from lowest to highest degree. The number of coefficients provided
determines the degree of the polynomial.

The `polynomial` object should have a field called `coefs` that stores the
coefficients as a vector, in order from lowest to highest degree.

For example, the call `poly(3, 1, 0, 2)` should construct a `polynomial` object
that represents the polynomial $3 + x + 2x^4$. The `coefs` field should contain
the vector `(3, 1, 0, 2)`.

If the `poly` function is called without arguments, or if any of the
coefficients are not numbers, the function should produce an error message
instead of returning a `polynomial` object.

_Hint 1: You'll need to use the `...` parameter here._

_Hint 2: You can convert a list to a vector with `unlist()` or `as.vector()`._

**YOUR ANSWER GOES HERE:**

```{r}
poly = function(...) {
  if (nargs() == 0 | !is.numeric(c(...))) {
    stop("Invalid arguments, must pass numeric coefs")
  }
  polynomial = list(coefs = c(...))
  structure(polynomial, class = "polynomial")
}
poly(3, 1, 0, 2)
```

S3 Generics & Methods
=====================

Watch the "S3 Generics & Methods" lecture video.


## Exercise 2

Create a `print` method for `polynomial` objects that prints them similar to
how you would write a polynomial mathematically. That is:

* Terms should be printed from lowest to highest degree.
* Every coefficient should be rounded to show at most 2 decimal places.
* It is okay to print terms where the coefficient is 0.

* The first term should just be a number. Only negative numbers should begin
  with a sign. Examples: `-0.67` and `2`.

* The second term should:
    + Begin with two spaces and a sign
    + End with `x`
    + Examples: `-0.67x` and `+2x`.

* The third term and all subsequent terms should:
    + Begin with two spaces and a sign
    + End with `x^d`, where `d` is the degree of the term
    + Examples: `-0.67x^2` and `+2x^4`

For example, the object created by `polynomial(2/3, 0, 1)` should print as
`0.67  +0x  -1x^2`.

Test your print method by for several different `polynomial` objects.

Your method should work even if `print` is not called explicitly. For example,
this code should implicitly call your `print` method:
```{r, eval = FALSE}
p = poly(1, 3, 2)
p
```

_Hint 1: Make sure your method's parameters match the generic `print`
function's parameters._

_Hint 2: Use vectorized operations to assemble the output string. Create a
prefix, coefficient, and suffix string for each term. Then use `paste` to paste
everything together._

_Hint 3: The `format` function can convert a number to a string. The `digits`,
`drop0trailing`, and `trim` parameters are especially useful for this
exercise._

_Hint 4: Until your `print` function is working correctly, you may want to give
it a different name. That way you can still see R's default printout for
`polynomial` objects while testing your function._

**YOUR ANSWER GOES HERE:**

```{r}
print.polynomial = function(polynomial) {
  output = ""
  power = 0
  for(coef in polynomial$coefs) {
    coef = round(coef, digits = 2)
    if(coef < 0) {
      if(power == 0) {
        output = paste(output, coef, "  ", sep = "")
      } else if (power == 1) {
        output = paste(output, coef, "x", "  ", sep = "")
      }else {
        output = paste(output, coef, "x^", power, "  ", sep = "")
      }
    } else if(power == 0) {
       output = paste(coef, "  ", sep = "")
    } else if (power == 1) {
      output = paste(output, "+", coef, "x", "  ", sep = "")
    } else {
      output = paste(output, "+", coef, "x^", power, "  ", sep = "")
    }
    power = power + 1
  }
  output = paste(output, "\n", sep = "")
  output = (output)
  cat(output)
}
p
print(poly(2/3, 0, 1))
print(poly(-1/6, -10, 42, 144, 169))
```

Other Object Systems
====================

Watch the "Other Object Systems" lecture video.


## Exercise 3

Read Chapter 16 of [Advanced R](https://adv-r.hadley.nz/).

Is threading state a problem in Homework 5, Exercise 1? Why or why not?

Based on this, is R6 strictly necessary to avoid the threading state problem?

_Hint: For more context, read Chapter 14 as well._

**YOUR ANSWER GOES HERE:**

No, Homework 5, Exercise 1 only requires us to create and return the stack object. R6 is not strictly necassary to avoid this problem.
