---
title: "STAT 33B Homework 7"
date: "Dec 10, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This homework is due __Dec 10, 2020__ by 11:59pm PT.

Homeworks are graded for correctness. 

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

If you have any last-minute trouble knitting, **DON'T PANIC**. Submit your Rmd
file on time and follow up in office hours or on Piazza to sort out the PDF.


SpamAssassin Email Data
=======================

The SpamAssassin Email Data set is a collection of email messages used to train
the SpamAssassin software to detect spam. The email messages are divided into
legitimate "ham" emails and illegitimate "spam" emails. Each email is in a
separate plain text file.

In this assignment, you'll only use a collection of "ham" emails. You can find
the emails in the file `emails.zip` on the bCourse. You will need to unzip the
file before proceeding with Exercise 1.

This data set is originally from the
[Apache SpamAssassin](https://spamassassin.apache.org/) project.


## Exercise 1

The `readLines` function reads lines of text from a file and returns them in a
character vector with one element for each line. The first argument is the path
to the file. By default, the function will read all of the lines in the file.

Write a function `read_email` that reads all of the text in a single email
file. Your function should have a parameter `file` to set the path to the file.
Your function should collapse all of the lines in the file into a single string
with lines separated by the newline character `\n`.

Show that your function works for 3 of the email files.

_Hint: The `paste` function is relevant here._

**YOUR ANSWER GOES HERE:**

```{r}
read_email = function(file) {
    lines = readLines(file)
    paste(lines, sep = "", collapse = "\n")
}

x = read_email("data/emails/easy_ham/0001.ea7e79d3153e7469e7a9c3e0af6a357e")
y = read_email("data/emails/easy_ham/0002.b3120c4bcbf3101e661161ee7efcb8bf")
z = read_email("data/emails/easy_ham/0003.acfc5ad94bbd27118a0d8685d18c89dd")
```

## Exercise 2

Write a function `read_email_all` that reads all of the files in the email
directory and returns a character vector with one element for each email. Your
function should have a parameter `dir` to set the path to the directory. Your
function should call the `read_email` function from Exercise 1.

Make sure not to put other files in the email directory!

After writing your function, use it to read all of the email files into a
character vector called `emails`.

How many email files are there?

_Hint: The `list.files` function is relevant here._

**YOUR ANSWER GOES HERE:**

```{r}
read_email_all = function(dir) {
    files = list.files(dir, full.names = TRUE)
    vapply(files, read_email, "a")
}

emails = read_email_all("data/emails/easy_ham")
length(emails)
```

## Exercise 3

Use stringr and regular expressions to write a function `extract_email_addr`
that extracts all email addresses from a character vector. Your function should
have a parameter `x` for the character vector, and should return a character
vector with one element for each email address (duplicates are okay).

For simplicity, you can assume the formatting rules for email addresses are
that they:

1. Must contain exactly one at-symbol `@`
2. Can also contain any number of letters, numbers, or characters in `._-`

Test your function on some made up strings and also on one of the email
messages.

_Hint: Using character classes `[ ]` in the regex pattern is important here._

_Note: It's not necessary for this exercise, but if you're curious about the
actual rules for email addresses, see
[Section 3.4.1 of RFC 5322](https://tools.ietf.org/html/rfc5322#section-3.4.1),
or [this Wikipedia article](https://en.wikipedia.org/wiki/Email_address)._

**YOUR ANSWER GOES HERE:**

```{r}
library(stringr)
extract_email_addr = function(x) {
    unlist(str_extract_all(x, "[[:alnum:]|.|_|-]+@[[:alnum:]|\\.|_|-]+"))
}
head(extract_email_addr(emails[1]))
```

## Exercise 4

Using your `extract_email_addr` function and the email message data:

1. How many different email addresses appear in the emails?
2. Which 5 email addresses appear the most?

**YOUR ANSWER GOES HERE:**

```{r}
addresses = unlist(sapply(emails, extract_email_addr), use.names = FALSE)
addresses_count = table(addresses)
# Number of unique addresses
length(addresses_count)

# Top 5 addresses
head(sort(addresses_count, decreasing = TRUE), 5)
```

## Exercise 5

The part of an email address after the `@` is called the _domain_. The domain
refers to a website, so it usually contains at least one dot. For example, Cal
email addresses use the domain `berkeley.edu`, which is also the Cal website
address.

Which 10 domains are the most common among the email addresses you extracted
from the email data?

How many domains in the email addresses end in `.edu`?

**YOUR ANSWER GOES HERE:**
```{r}
domains = substring(str_extract(addresses, "@.+"), 2)
domains_count = sort(table(domains), decreasing = TRUE)
# 10 most common domains
head(domains_count, 10)

edu_domains = domains[str_detect(domains, "\\.edu$")]
# Number of .edu domains
edu_domains_count = table(edu_domains)
length(edu_domains_count)
```
