---
title: "Stat 33B - Lecture Notes 4"
date: September 13, 2020
output: pdf_document
---


Three Ways to Subset
====================

Use the subset (or "square bracket") operator `[` to get elements from a vector
by position:
```{r}
x = rep(c(4.1, 2.2, 7.9), 2)
x

x[3]

x[c(5, 4, 5)]
```

You can also use the subset operator to set elements:
```{r}
x[1] = 10

x

x[c(4, 5)] = c(-20, -10)
x
```

Negative positions mean "everything except":
```{r}
x[-1]

x[-c(5, 6)]

# x[c(5, -1)]
```


The subset operator `[` can actually get/set elements in three ways:

* By position
* By name
* By condition


## Subsets by Name

You can make vectors with named elements (just like lists):
```{r}
x = c(a = 6, b = 4, c = 3)
x

names(x)

names(x) = c("hi", "hello", "goodbye")
x
```

You can use the subset operator `[` to get elements by name:
```{r}
x["hello"]

x[c('hi', 'hi', 'goodbye')]
```

Likewise to set elements by name:
```{r}
x["hello"] = 21

x
```



## Congruent Vectors

Two vectors are **congruent** if they have the same length and they correspond
elementwise.

For example, suppose you do a survey that records each person's:

* Favorite animal
* Age

These are two different vectors of information, but each person will have a
response for both.

So you'll have two vectors that are the same length, with corresponding
elements:
```{r}
animal = c("dog", "cat", "iguana")
age = c(31, 24, 72)
```

These vectors are congruent.

Columns in a data frame are always congruent!



## Subsets by Condition

The third way to subset a vector with `[` is to use a congruent logical vector.

For example:
```{r}
x = c(2.2, 3.1, 6.7)
x

logic = c(TRUE, FALSE, TRUE)

x[logic]
```

`TRUE` means keep the element, `FALSE` means drop the element.

Missing values `NA` in the logical vector are retained in the result:
```{r}
x[c(TRUE, NA, NA)]
```


A **condition** is any expression that returns a logical vector.

For example, comparisons are conditions:
```{r}
x > 3
```

You can use conditions to take subsets:
```{r}
x[x > 3]
```


Final note: you can technically subset with any logical vector.

If the length is different, R uses the recycling rule:
```{r}
x[c(TRUE, FALSE)] #TRUE
```






Logic
=====

R provides functions to negate, combine, and summarize logical vectors.



## Logical Negation

The **NOT** operator `!` converts `TRUE` to `FALSE` and `FALSE` to `TRUE`:
```{r}
x = c(TRUE, FALSE, TRUE, TRUE, NA)

x

!x
```

You can use `!` with a condition:
```{r}
y = c("hi", "hello")
!(y == "hi")
```

The NOT operator is vectorized.



## Logical Combinations

R also has operators for combining logical values.

The **AND** operator `&` returns `TRUE` only when both arguments are `TRUE`.

For example:
```{r}
FALSE & FALSE
TRUE & FALSE
FALSE & TRUE
TRUE & TRUE

c(TRUE, FALSE, TRUE) & c(TRUE, TRUE, FALSE)
```

The **OR** operator `|` returns `TRUE` when at least one argument is `TRUE`.

For example:
```{r}
FALSE | FALSE
TRUE | FALSE
FALSE | TRUE
TRUE | TRUE

c(TRUE, FALSE) | c(TRUE, TRUE)
```

Be careful: everyday English is less precise than logic.

You might say:

> I want all subjects with age over 50 and all subjects that like cats.

But in logic this means:

`(subject age over 50) OR (subject likes cats)`

So think carefully about whether you need both conditions to be true (AND) or
at least one (OR).


Rarely, you might want _exactly one_ condition to be true.

The **XOR** (eXclusive OR) function `xor()` returns `TRUE` when exactly one
argument is `TRUE`.

For example:
```{r}
xor(FALSE, FALSE)
xor(TRUE, FALSE)
xor(TRUE, TRUE)
```

The AND, OR, and XOR operators are vectorized.



## Short-circuit Operators

The second argument is irrelevant in some conditions:

* `FALSE &` is always `FALSE`
* `TRUE |` is always `TRUE`

Now imagine you have `FALSE & long_computation()`. You can save time by
skipping `long_computation()`.

A **short-circuit** operator does exactly that.


R has two:

* `&&` is a short-circuited `&`
* `||` is a short-circuited `|`

These operators only evaluate the second argument if it is necessary to
determine the result.

These operators are _not_ vectorized!

Examples:
```{r}
TRUE && FALSE
TRUE && TRUE

TRUE || TRUE

c(TRUE, FALSE) && c(TRUE, TRUE)
```






Logical Summaries
=================

Use the `any()` function to check that a logical vector contains a `TRUE`
value:
```{r}
x = c(TRUE, FALSE, TRUE, TRUE, TRUE)

any(x)

any(c(FALSE, FALSE, NA))

any(c(TRUE, FALSE, NA))
```

Use the `all()` function to check that a logical vector contains only `TRUE`
values:
```{r}
all(c(TRUE, TRUE, FALSE))

all(c(TRUE, TRUE))

all(c(TRUE, TRUE, NA))
```

Use `sum()` or `table()` to count `TRUE` values:
```{r}
sum(x)

table(x)
```

The `sum()` function is slightly faster than `table()`.



## The `which()` Function

Sometimes you'll want to ignore missing values in a condition:
```{r}
x = c(5.1, NA, 3.2)
x > 4

x[x > 4]
```

The `which()` function returns the position of each `TRUE` in a logical vector:
```{r}
which(x > 4)
```

The function ignores missing values.

So you can use `which()` to ignore missing values in a condition:
```{r}
x[which(x > 4)]
```

Do this as needed, not everywhere.

Calling a function takes more time than not calling a function!









Subset Versus Extract
=====================

We've learned two different ways to get/set elements:

1. `[`, the subset operator ("square bracket")

2. `[[`,  the extraction operator ("double square bracket")

Both work with vectors, arrays, lists, data frames, and more.

That said, they aren't identical.



## Multiple Elements

The subset operator `[` can get/set any number of elements:
```{r}
x = c(5, 8.1, 2, 10)

x[c(1, 1, 3)]
```

The extraction operator `[[` can only get/set a single element:
```{r}
x[[c(1, 2)]]

x[[1]]

x[[4]] = 21
x

x[[c(2, 3)]] = 7
```

Note that a vector can be a *single element* of a list:
```{r}
li = list(c(1, 2, 3), "hello")
li

li[[1]]
```

You can't use `[[` to get/set elements by condition, because the logical vector
might have more than one `TRUE` element:
```{r}
x[[c(T, F, F, F)]]
```

Both `[` and `[[` can get/set elements by position and by name.

```{r}
x = c(a = 1, b = 2)

x[["a"]]
```


## Invalid Elements

The subset operator `[` returns `NA` for invalid elements:
```{r}
x[3]
```

The extraction operator `[[` raises an error for invalid elements:
```{r}
x[[3]]
```

## Containers

In programming languages, a **container** is an object that contains other
objects.


Lists are R's primary container.

In lists, elements can have different data types:
```{r}
x = list(c(1, 2, 3), sin, c("hi", "hello"))
```


The subset operator `[` keeps containers:
```{r}
x[1]
class(x[1])
```

The extraction operator `[[` drops containers:
```{r}
x[[1]]
class(x[[1]])
```













Subsets of Data Frames
======================

Let's look at the dogs data again:
```{r}
dogs = readRDS("data/dogs/dogs.rds")
dogs
```

R thinks of data frames as lists of columns.

So one-dimensional `[` and `[[` get columns:
```{r}
class(dogs[1])

class(dogs[[1]])
```
`[` keeps the data frame, `[[` drops the data frame.



## Two-dimensional Subsetting

Data frames are two-dimensional, so you can also use matrix-style subsetting.

That is, use the syntax:

    DATA[ROWS, COLUMNS]

This works for the subset operator `[` and the extraction operator `[[`.

Examples:
```{r}
class(dogs[1:3, 2])

dogs[[1, 2]]

dogs[1:3, ]

dogs[2:4, -2]

dogs[1:3, c("popularity", "breed")]
```



## The Drop Parameter

If you use two-dimensional subsetting with `[` to select exactly one column,
you get a vector:
```{r}
dogs[1:3, 2]
```

The container is dropped!

This is also true for matrices:
```{r}
x = matrix(1, 2, 2)
x
x[, 1]
```


You can control this behavior with the `drop` parameter:
```{r}
x[, 1, drop = FALSE]

dogs[1:3, 2, drop = FALSE]
```

The default is `drop = TRUE`.



## The Dollar Sign

The `$` operator is similar to `[[`, but only gets elements by name:
```{r}
dogs$breed
dogs[["breed"]]
```

Unlike `[[`, the `$` will try to match partial names.

```{r}
dogs$bree
```


The `$` allows quotes around the name:
```{r}
dogs$"breed"
```
Use quotes for names with non-alphanumeric characters.



## The `subset()` Function

Suppose we want to get rows of a data frame that satisfy a condition:
```{r}
dogs[which(dogs$weight > 50), ]
```

The `subset()` function is a shortcut to avoid typing the name of the data
frame over and over when subsetting rows with a condition:
```{r}
subset(dogs, weight > 50)
```
With `subset()`, the data frame the columns come from is implied.



The `subset()` function also eliminates `NA` values in the condition:
```{r}

```
