---
title: "STAT 33B Workbook 3"
date: "Sep 17, 2020"
author: "YOUR NAME (YOUR SID)"
output: pdf_document
---

This workbook is due __Sep 17, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.


Three Ways to Subset
====================

Watch the "Three Ways to Subset" lecture video.


## Exercise 1

Create a variable `count` that contains the integers from 1 to 100 (inclusive).

The `as.character()` function coerces its argument into a character vector.
Coerce `count` into a character vector and assign the result to a variable
called `fizzy`. Now you have congruent vectors `count` and `fizzy`.

Use subset assignment to replace every number in `fizzy` that's:

* Divisible by 3 with `"Fizz"`
* Divisible by 5 with `"Buzz"`
* Divisible by 15 with `"FizzBuzz"`

Leave all other numbers in `fizzy` as-is.

Print out the final version of `fizzy`. It should begin:
```
 [1] "1"        "2"        "Fizz"     "4"        "Buzz"     "Fizz"
 [7] "7"        "8"        "Fizz"     "Buzz"     "11"       "Fizz"
[13] "13"       "14"       "FizzBuzz" "16"       "17"       "Fizz"
```

_Hint 1: Take advantage of the fact that `count` and `fizzy` are congruent._

_Hint 2: The modulo operator `%%` returns the remainder after dividing its
first argument by its second argument. You can use the modulo operator to test
whether a number is divisible by some other number (that is, the remainder is
zero after divsion)._


YOUR ANSWER GOES HERE:



Logic
=====

Watch the "Logic" lecture video.


## Exercise 2

Suppose you conduct a survey and store the results in the following congruent
vectors:

```{r}
# Q: What's your favorite color?
color = c("red", "blue", "blue", "green", "yellow", "green")
color = factor(color)

# Q: Name a dessert you like?
sweet = c("egg tart", "brownie", "ice cream", "ice cream", "fruit", "egg tart")
sweet = factor(sweet)

# Q: Name a desert (not dessert) you like?
dry = c("Kalahari", "Atacama", "Taklamakan", "Sonoran", "Atacama", "Atacama")
dry = factor(dry)

# Q: How old are you?
age = c(23, 15, 92, 21, 28, 45)

# Q: How many UFOs have you seen since 2010?
ufo = c(0, 3, 122, 0, 0, 1)
```

Use the vectors above, comparison operators, and logical operators to compute a
logical vector that corresponds to each of the following conditions.

1. People who have seen a UFO.

2. People who have seen a UFO but aren't over 50 years old.

3. People who didn't choose ice cream.

4. People who like both ice cream and the color green.

5. People who like the color red or the color green.

YOUR ANSWER GOES HERE:



## Exercise 3

In the expression `(x < 5) == TRUE`, explain why `== TRUE` is redundant.

YOUR ANSWER GOES HERE:



Logical Summaries
=================

Watch the "Logical Summaries" lecture video.


No exercises for this section. You're halfway finished!



Subset vs. Extract
==================

Watch the "Subset vs. Extract" lecture video.


## Exercise 4

A **recursive** list is a list with elements that are also lists.

Here's an example of a recursive list:
```{r}
mylist = list(list(1i, 2, 3i), list(c("hello", "hi"), 42))
```

Use the recursive list above to answer the following:

1. What's the first element? What's the second element?

2. Use the extraction operator `[[` to get the value `3i`.

3. What does `mylist[[c(1, 3)]]` do? What does the index `c(1, 3)` mean here?
   Experiment with using other vectors in `[[` to get elements from the
   recursive list. Then explain what the extraction operator `[[` does for
   recursive lists when the index is a vector.

YOUR ANSWER GOES HERE:


## Exercise 5

For the list `cool_list = list("Hope", "springs", "eternal")`, why is
`cool_list[1]` the same as `cool_list[1][1][1][1][1]`? Is this property unique
to `cool_list`, or is it a property of all lists? Explain your answer.

YOUR ANSWER GOES HERE:



Subsets of Data Frames
======================

Watch the "Subsets of Data Frames" lecture video.

## Exercise 6

For the dogs data, compute:

1. The subset that contains rows 10-20 of the height, weight, and longevity
   columns.

2. The mean and median of the longevity column (ignoring missing values).

3. The number of dog breeds whose average weight is greater than 42. _Note: the
   `weight` column is the average weight of each row's breed._

4. The subset of large dogs that require daily grooming.

YOUR ANSWER GOES HERE:



## Exercise 7

The `sort()` function sorts the elements of a vector. For instance:
```{r}
x = c(4, 5, 1)
sort(x)
```

The `order()` function is a more flexible alternative to `sort()`. Instead of
returning the sorted vector, the `order()` function returns the index that
sorts the vector. To actually sort the vector, you have to pass this index to
the subset operator `[`:
```{r}
x = c(4, 5, 1)
x[order(x)]
```

The advantage of `order()` over `sort()` is that you can use `order()` to sort
one vector based on the elements of some other congruent vector.

Use the `order()` function to sort the rows of the dogs data set based on
height. What are the 3 tallest breeds of dog?

YOUR ANSWER GOES HERE:
