---
title: "Stat 33B - Lecture Notes 12"
date: November 8, 2020
output: pdf_document
---

Code Performance
================

See the slides.



Profiling
=========

The `Rprof` function enables or disables R's built-in profiler.

Important parameters:

* `filename`, a path to a file for storing results. Defaults to `Rprof.out`
* `interval`, the time between samples, in seconds
* `memory.profiling`, whether to track memory in addition to time

Call `Rprof(NULL)` to disable the profiler.


As an example, let's look at what happens if we don't preallocate.

We'll use the rejection sampler from Homework 4:
```{r}
dhand = function(x) {
  y = numeric(length(x))
  i = -1.6 < x & x < 1.6
  y[i] = sqrt(2.56 - x[i]^2) + dunif(x[i], -1.25, -0.75) +
    dunif(x[i], -0.5, 0.5)

  y
}


rhand = function(n) {
  samp = c()

  while (n > 0) {
    x = runif(1, -1.6, 1.6)
    y = runif(1, 0, 3.5)

    if (y < dhand(x)) {
      samp = c(samp, x)
      n = n - 1
    }
  }

  samp
}
```

```{r}

```

The `summaryRprof` function summarizes results from `Rprof`.

Important parameters:

* `filename`, the path to the results file. Defaults to `Rprof.out`
* `memory`, how to display memory information. Use `"both"` to see total
  changes.

The summary lists times in seconds and memory in bytes.

Let's try it:
```{r}

```

The `total.pct` column is the time spent in a function as a percentage of
total time.

The `self.time` column is the time spent in a function and NOT other functions
called from that function.

The `self.pct` column is the `self.time` as a percentage of total time.

The `by.self` data frame only includes rows where `self.time` is nonzero.



Profvis
=======

The profvis package provides a visualization of profiler results.

Load the library and then call `profvis` on the code you want to profile.

Under the hood, profvis uses the `Rprof` function.


Let's again look at the rejection sampler from Homework 4.

This time we'll use profvis:
```{r}

```



Profiling Case Study
====================

In a 2-dimensional random walk, at each step we randomly choose a move in one
of the four compass directions.

Goal: write an efficient function to compute the path of a 2-d random walk.

```{r}

```
