---
title: "Stat 33B - Lecture Notes 7"
date: October 4, 2020
output: pdf_document
---


Functions
=========

You can create functions with the `function` command.

Functions return the value of their final expression:
```{r}

```
Indent the body of the function by 2 or 4 spaces.



Curly braces are optional if the body is a single expression:
```{r}

```


To exit the function and return a result early, use `return()`:
```{r}

```
It's idiomatic to only use `return()` when strictly necessary.



Default Arguments
=================

Parameters can have default arguments that are used if no argument is provided.

Use `=` (not `<-`) to assign default arguments:
```{r}

```



Function Example
================

```{r}

```

How many heatwaves were there? (periods of days where TMAX >= 100)

First: days TMAX >= 100
```{r}

```

Simple example:
```{r}

```

Turn this into a function:
```{r}

```

Try it on on TMAX:
```{r}

```



Variables: Scope & Lookup
=======================

Local Variables
---------------

A variable's *scope* is the section of code where it exists and is accessible.

The `exists()` function checks whether a variable is in scope:
```{r}

```

When you create a function, you create a new scope.

Variables defined inside of a function are *local* to the function.


Local variables cannot be accessed from outside:
```{r}

```

Local variables are reset each time the function is called:
```{r}

```


The `local()` function also creates a new scope:
```{r}

```
Rarely used in practice.



Lexical Scoping
---------------

A function can use variables defined outside (non-local), but only if those
variables are in scope where the function was **defined**.

This property is called *lexical scoping*.

For example:
```{r}

```

Variables defined directly in the R console are *global* and available to any
function.


Local variables *mask* (hide) non-local variables with the same name:
```{r}

```


ONE EXCEPTION!

We often call variables:
```{r}

```

In this case, the variable must refer to a function, so R ignores local
variables that aren't.

For example:
```{r}

```


Dynamic Lookup
--------------

Variable lookup happens when a function is **called**, not when it's defined.

This is called *dynamic lookup*.

For example:
```{r}

```


Summary
-------

* Function definitions (or `local()`) create a new scope.

* Local variables
    + Are private
    + Get reset for each call
    + Mask non-local variables (exception: function calls)

* *Lexical scoping*: where a function is **defined** determines which non-local
  variables are in scope.

* *Dynamic lookup*: when a function is **called** determines values of
  non-local variables.









Lazy Evaluation
===============

Arguments to a function are only evaluated if they're used:
```{r}

```
This is called *lazy evaluation*.



You can use one parameter in the default argument of another:
```{r}

```
Often used to provide multiple ways to call a function.



You can force evaluation with the `force()` function:
```{r}

```
Rarely needed in practice.










The Dots Parameter
==================

The *dots* parameter `...` accepts any number of arguments.



The dots parameter is often used to forward arguments to another function.

For example:
```{r}

```


You can access elements of `...` with the `...elt()` function:
```{r}

```
This only evaluates the argument you accessed.


You can get the number of arguments with `...length()`:
```{r}

```
This doesn't evaluate the arguments.


You can convert `...` to a list with the `list()` function:
```{r}

```
This evaluates all of the arguments.



See `?dots` for more about the dots parameter.



Using Functions
===============

Functions are the building blocks for solving problems.

Take a divide-and-conquer approach, breaking large problems into
smaller steps.

Use a short function for each step. Then it's easier to:

* Test that each step works correctly.
* Modify, reuse, or repurpose a step.

Before you write a function, it's useful to go through several steps:

1. Write down what you want to do, in detail. It can also help to
   draw a picture of what needs to happen.

2. Check whether there's already a built-in function. Google it!

3. Write the code to handle a simple case first. For data science
   problems, use a small dataset at this step.
