---
title: "Stat 33B - Lecture Notes 7"
date: October 4, 2020
output: pdf_document
---


Functions
=========

You can create functions with the `function` command.

Functions return the value of their final expression:
```{r}
square = function(x) {
  x^2
}
```
Indent the body of the function by 2 or 4 spaces.



Curly braces are optional if the body is a single expression:
```{r}
square = function(x) x^2
```


To exit the function and return a result early, use `return()`:
```{r}
square = function(x) {
  if (!is.numeric(x))
    return (NA)
  
  x^2
}

square(6)
square("hi")

square = function(x) {
  if (!is.numeric(x)) NA else x^2
}

square(3)
square("test")
```
It's idiomatic to only use `return()` when strictly necessary.











Default Arguments
=================

Parameters can have default arguments that are used if no argument is provided.

Use `=` (not `<-`) to assign default arguments:
```{r}
# Gets the top n values from a vector
top = function(x, n = 5) {
  sorted = sort(x, decreasing = TRUE)
  head(sorted, n)
}

y = c(-6, 7, 10, 3, 1, 15, -2)
top(y, 3)

top(y)
```













Function Example
================

```{r}
w = read.csv("data/weather/weather_daily.csv")

head(w)
```

How many heatwaves were there? (periods of days where TMAX >= 100)

First: days TMAX >= 100
```{r}
sum(w$TMAX >= 100)
```

Simple example:
```{r}
# 2 heatwaves
temp = c(95, 96, 98, 101, 105, 114, 102, 93, 91, 103, 104, 87)
#temp = c(94, 100, 93, 102, 92, 105, 86)

is_hot = temp >= 100

change = c(FALSE, tail(is_hot, -1) != head(is_hot, -1))
change
num_changes = sum(change)
if (num_changes %% 2 == 0) {
  num_changes / 2
} else {
  (num_changes + 1) / 2
}
```

Turn this into a function:
```{r}
count_heatwaves = function(temp) {
  is_hot = temp >= 100

  change = c(FALSE, tail(is_hot, -1) != head(is_hot, -1))
  num_changes = sum(change)
  if (num_changes %% 2 == 0) {
    num_changes / 2
  } else {
    (num_changes + 1) / 2
  }
}

count_heatwaves(temp)

count_heatwaves(c(10, 100, 10, 100, 100))

count_heatwaves(c(100, 10, 100, 100, 10))

# Here's a bug:
count_heatwaves(c(100, 10, 100, 100))
```

Try it on on TMAX:
```{r}
head(w$TMAX)

count_heatwaves(w$TMAX)
```








Variables: Scope & Lookup
=======================

Local Variables
---------------

A variable's *scope* is the section of code where it exists and is accessible.

The `exists()` function checks whether a variable is in scope:
```{r}
exists("x")

x = 3
exists("x")
```

When you create a function, you create a new scope.

Variables defined inside of a function are *local* to the function.


Local variables cannot be accessed from outside:
```{r}
rescale = function(x, center, scale) {
  centered = x - center
  centered / scale
}

centered
exists("centered")
```

Local variables are reset each time the function is called:
```{r}
f = function() {
  is_z_in_scope = exists("z")
  z = 42
  
  is_z_in_scope
}

f()
f()
```


The `local()` function also creates a new scope:
```{r}
local({
  hi = 10
})

exists("hi")
```
Rarely used in practice.



Lexical Scoping
---------------

A function can use variables defined outside (non-local), but only if those
variables are in scope where the function was **defined**.

This property is called *lexical scoping*.

For example:
```{r}
cats = 3
get_cats = function() cats

get_cats()


create_dogs = function() {
  dogs = "hello"
}

get_dogs = function() dogs

create_dogs()
get_dogs()
```

Variables defined directly in the R console are *global* and available to any
function.


Local variables *mask* (hide) non-local variables with the same name:
```{r}
get_parrot = function() {
  parrot = 3
  
  parrot
}

parrot = 42
get_parrot()
```


ONE EXCEPTION!

We often call variables:
```{r}
#mean()
```

In this case, the variable must refer to a function, so R ignores local
variables that aren't.

For example:
```{r}
my_mean = function() {
  mean = 0
  
  mean(c(1, 2, 3))
}

my_mean()

my_get_cats = function() {
  get_cats = 10
  
  get_cats()
}

my_get_cats()
```


Dynamic Lookup
--------------

Variable lookup happens when a function is **called**, not when it's defined.

This is called *dynamic lookup*.

For example:
```{r}
cats = 10
get_cats()

cats = 20
get_cats()
```


Summary
-------

* Function definitions (or `local()`) create a new scope.

* Local variables
    + Are private
    + Get reset for each call
    + Mask non-local variables (exception: function calls)

* *Lexical scoping*: where a function is **defined** determines which non-local
  variables are in scope.

* *Dynamic lookup*: when a function is **called** determines values of
  non-local variables.









Lazy Evaluation
===============

Arguments to a function are only evaluated if they're used:
```{r}
f = function(x) 42

f(6)

f()

message("hi")

f(message("hello"))
```
This is called *lazy evaluation*.



You can use one parameter in the default argument of another:
```{r}
add = function(x = y, y = x) x + y

add(x = 4)

add(y = 3)

add()
```
Often used to provide multiple ways to call a function.



You can force evaluation with the `force()` function:
```{r}
f = function(x) {
  force(x)
  42
}

f()

f(message("hi"))
```
Rarely needed in practice.










The Dots Parameter
==================

The *dots* parameter `...` accepts any number of arguments.



The dots parameter is often used to forward arguments to another function.

For example:
```{r}
# Mean function with tolerance:
mean_tol = function(x, tol, ...) {
  mean(x[x > tol], ...)
}

mean_tol(c(1, 3, 5, 0.01, 0.2, NA), 0.5)

mean_tol(c(1, 3, 5, 0.01, 0.2, NA), 0.5, na.rm = TRUE)

mean(c(1, 3, 5, 10))
```


You can access elements of `...` with the `...elt()` function:
```{r}
f = function(x, ...) {
  ...elt(2)
  x + ...elt(1)
}

f(3, 5, message("hi"))

f(3)
```
This only evaluates the argument you accessed.


You can get the number of arguments with `...length()`:
```{r}
count_args = function(...) ...length()

count_args(4, 5, 6, message("hi"))
```
This doesn't evaluate the arguments.


You can convert `...` to a list with the `list()` function:
```{r}
f = function(...) list(...)

f(hi = 1, 3, 4)
```
This evaluates all of the arguments.



See `?dots` for more about the dots parameter.



Using Functions
===============

Functions are the building blocks for solving problems.

Take a divide-and-conquer approach, breaking large problems into
smaller steps.

Use a short function for each step. Then it's easier to:

* Test that each step works correctly.
* Modify, reuse, or repurpose a step.

Before you write a function, it's useful to go through several steps:

1. Write down what you want to do, in detail. It can also help to
   draw a picture of what needs to happen.

2. Check whether there's already a built-in function. Google it!

3. Write the code to handle a simple case first. For data science
   problems, use a small dataset at this step.
