---
title: "Stat 33B - Lecture Notes 6"
date: September 27, 2020
output: pdf_document
---


Apply Function Basics
=====================

Doing the same operation repeatedly is a common pattern in programming.

Vectorization is one way, but not all functions are vectorized.


In R, the "apply functions" are another way to do something repeatedly.

The apply functions call a function on each element of a vector or list.



The `lapply()` Function
---------------------

The first and most important apply function is `lapply()`. The syntax is:
```
lapply(X, FUN, ...)
```

The function `FUN` is called once for each element of `X`, with the element as
the first argument. The `...` is for additional arguments to `FUN`, which are
held constant across all calls.


Unrealistic example:
```{r}

```
In practice, it's clearer and more efficient to use vectorization here.


Let's use the dogs data for some realistic examples:
```{r}

```

`lapply()` always returns the result as a list.

"l" for **list** result.



The `sapply()` Function
---------------------

`sapply()` simplifies the result to a vector, when possible.

"s" for **simplified** result.

Examples:
```{r}

```

The `sapply()` function is useful if you are working interactively.















The Split-Apply Strategy
========================

The `split()` function splits a vector or data frame into groups based on some
other vector (usually congruent).

```{r}

```


Split weight of dogs by the group column:
```{r}

```

The `split()` function is especially useful when combined with `lapply()` or
`sapply`().

```{r}

```
This is an R idiom!



The `tapply()` Function
---------------------

The `tapply()` function is equivalent to the `split()` and `sapply()` idiom.

"t" for **table**, because `tapply()` is a generalization of the
frequency-counting function `table()`.


Examples:
```{r}

```

This strategy is important for analyzing tabular data regardless of what
programming language or packages you're using.
















Even More Apply Functions
=========================


The `vapply()` Function
---------------------

`vapply()` simplifies the result to a vector of a specific data type.

"v" for **vector** (or matrix) result.


Examples:
```{r}

```

The `vapply()` function is more robust and efficient than `sapply()` because
the return type is specified.

Use `vapply()` when you write functions or other non-interactive code.



The `mapply()` and `Map()` Functions
--------------------------------

`mapply()` applies a function to multiple data arguments.

"m" for **multiple** arguments.

Examples:
```{r}

```

`mapply()` simplifies the result to a vector, when possible.

`Map()` is a wrapper for `mapply()` that never simplifies the result.



The `apply()` Function
--------------------

What if we want to compute the standard deviation of each row in a matrix?

The `apply()` function applies a function along one or more dimensions of an
array. The syntax is:
```
apply(X, MARGIN, FUN, ...)
```
The `MARGIN` is a vector of dimensions to apply the function over. 1 means
rows, 2 means columns, and so on.


For example:
```{r}

```
`apply()` returns a vector or matrix result.


`apply()` is rarely used, and when it is, it's usually for matrices.















Choosing an Apply Function
==========================


1. Is the function you want to apply vectorized?

   If yes, use vectorization.

   Otherwise, continue to #2.


2. Do you want to apply the function to elements or to groups?

   For elements, continue to #3.

   For groups, use the split-apply pattern. Use `split()`, then
   continue to #3 to choose an apply function.

   Note `tapply()` is equivalent to `split()` and `sapply()`.


3. Will the function return the same data type for each element?

   If yes, continue to #4.

   Otherwise, use `lapply()`.


4. Are you working interactively?

   If yes, use `sapply()`.

   Otherwise, use `vapply()`.


Other Apply Functions
---------------------

See this StackOverflow Post for a summary:

    https://stackoverflow.com/a/7141669


The purrr and dplyr packages provide Tidyverse alternatives to apply functions.







Conditional Expressions
=======================

The syntax for an if-statement in R is:
```{r}

```

The `else` block is optional:
```{r}

```

Curly braces `{ }` are optional for single-line expressions:
```{r}

```

The condition has to be a scalar:
```{r}

```

You can chain together if-statements:
```{r}

```

If-statements return the value of the last expression in the evaluated block:
```{r}

```

When there's no `else` block, the value of the `else` block is `NULL`:
```{r}

```















The `switch()` function
=====================

The `switch()` function uses integer or string matching to select an expression
to evaluate.

String example:
```{r}

```

Integer example:
```{r}

```

`switch()` only evaluates the selected expression.

So `switch()` is more efficient than using a list and subsetting:
```{r}

```






















The Congruent Vectors Strategy
==============================

If-statements don't work well with vectors.

For example, suppose we want to transform a vector `x` so that:

* Negative elements are set to 0.
* Positive elements are squared.

Using an if-statement doesn't work for this:
```{r}

```


Instead, use congruent vectors:

1. An input vector (or vectors) to use in conditions.

2. An output vector to store the results.

Use the input vector to conditionally assign elements to the output vector.


So:
```{r}

```


Another example:
```{r}

```


The `ifelse()` Function
-----------------------

R also has a vectorized `ifelse()` function.

For example:
```{r}

```

The `ifelse()` function is less efficient than a regular if-statement or the
congruent vectors strategy.
