---
title: "Stat 33B - Lecture Notes 9"
date: October 18, 2020
output: pdf_document
---


Printing Output
===============

The `cat` function prints a message to the console:
```{r}
cat("Hello")

cat(4)

cat("Hello", "Nick")
```

Strings lose their quotes and escape sequences are converted.

The `cat` function always returns `NULL`:
```{r}
x = cat("Hello")
x
```


The `print` function prints R's representation of an object to the console:
```{r}
print(5)

print(c(5, 6))

x = c(5, 6)
print(x)

print("Hello")
```

Strings keep their quotes and escape sequences are not converted.

The `print` function returns the object being printed:
```{r}
y = print(5)
y
```

The `show` function is similar to `print`, but always returns `NULL`.

A later lecture might cover why both `print` and `show` are necessary.


For most output, `print` is the wrong function to use!



Formatting Output
-----------------

_Escape sequences_ allow you to write characters that aren't on your keyboard
in strings.

Escape sequences always begin with a backslash.

See `?Quotes` for a complete list.

For example, `\n` is a newline:
```{r}
x = "Hello\nNick"

cat(x)

print(x)
```

The `paste` function combines strings:
```{r}
y = paste("Hello", "Nick")
y
```

The `sprintf` function can substitute values into strings.

Substitutions are marked with `%` followed by a character:
```{r}
sprintf("My age is %i, and my name is %s", 32, "Nick")
```



Messages, Warnings, and Errors
==============================

The `message`, `warning`, and `stop` functions form R's _condition system_.

The condition system provides a way to report and handle unusual conditions.



The `message` function prints message to the R console:
```{r}
message("Hello")
```


The `warning` function prints a warning:
```{r}
if (c(TRUE, FALSE)) 42

warning("Hi")
```

By default, calling `warning` in a function prints the name of the function:
```{r}
f = function(x, y) {
  warning("This is a warning!")
  x + y
}

f(3, 4)
```

You can use the `call.` parameter to control this:
```{r}
f = function(x, y) {
  warning("This is a warning!", call. = FALSE)
  x + y
}

f(3, 4)
```


The `stop` function prints an error message and stops evaluation:
```{r}
stop("Hi")

f = function(x, y) {
  stop("This is an error!")
  x + y
}

f(3, 4)
```
The `stop` function also has a `call.` parameter.



Handling Warnings and Errors
============================

Use the `try` function to try running an expression that might produce an
error:
```{r}
x = try(5 + 6)

y = try(5 + "hi")

if (inherits(y, "try-error")) {
  # Handle error
} else {
  # Proceed normally
}

```

If the expression produces an error, `try` returns an object with class
`try-error`. Otherwise, it returns the result.

An error in `try` does NOT stop evaluation:
```{r}
f = function(x, y) {
  try(5 + "hi")
  x + y
}
```


The error can be silenced by setting `silent = TRUE`:
```{r}
f = function(x, y) {
  try(5 + "hi", silent = TRUE)
  x + y
}
```

The `tryCatch` function provides another way to handle conditions.

The function takes "handlers" for different conditions:

* `message`
* `warning`
* `error`
* `interrupt` -- when the user interrupts the code (for example, by
  pressing Ctrl-C)

Each handler should be a function with one parameter.

If the expression produces a message/warning/error, `tryCatch` calls the
handler with information about the message/warning/error, and returns the
result. Otherwise, it returns the result of the expression.

For example:
```{r}
f = function(x, y) {
  #message("Hi")
  x + y
}

g = function(x, y) {
  stop("Hi")
  x + y
}

msg = tryCatch(f(3, 4), message = function(e) e)

err = tryCatch(g(3, 4), error = function(e) e)
```

`tryCatch` always silences the message/warning/error.



Global Options
==============

The `options` function sets global options for the R session.

Options have many uses (see `?options`).

One is controlling errors and warnings.



Example: the `warn` option controls how warnings R displayed.

To make R treat warnings as errors:
```{r}
f = function(x , y) {
  warning("Hi")
  x + y
}
```
This is useful for finding the source of a warning.


The `warn` option can be set to three different values:

* `warn = 0` (the default) warnings are only displayed after code finishes
  running

* `warn = 1` warnings are displayed immediately

* `warn = 2` warnings stop code from running, like errors



It's possible to override the default options every time R starts.

At startup, R runs code from an `.Rprofile` file, usually located in
your system's home directory.

See `?Startup` for full details.



A few other useful options to make R warn about partial name matches in calls
and the `$` operator:

* `warnPartialMatchAttr = TRUE`

* `warnPartialMatchDollar = TRUE`

* `warnPartialMatchArgs = TRUE`

For example:
```{r}
x = list(hello = 4)
x$he
```



Preventing Bugs
===============

Modularity
----------

Try to break problems into smaller steps.

Write down the input(s) and output(s) for each step.

Create functions for steps that are reusable. Use parameters for input and the
return value for output.

Be wary of extremely short (1 line) or long (> 20 line) functions. Often a sign
of failure to break down the problem.



Code Clarity
------------

Format your code so that it's easy to read:

* Use whitespace:

    + Put spaces after commas and around binary operators.
    + Indent code inside curly braces `{ }` by 2 spaces or 1 tab.
    + Don't mix tabs and spaces.
    + Separate logical steps or "paragraphs" with blank lines.

* Put closing curly braces `}` on their own line.
    + Exception: use `} else {`

* Use a consistent naming style:
   + `lowercase_with_underscores`
   + `camelCase`

* Use descriptive variable names. Short names are okay for frequently-used
  variables where the context makes the meaning apparent.

* Use comments:
    + To create a big picture plan for how to write your code.
    + To explain tricky code.
    + To summarize the purpose of a "paragraph" of code.
    + To document how to use your functions. Also see roxygen2.


Without formatting:
```{r}
f=function(x,xx){
xx=match.arg(xx,c("celsius","fahrenheit"),several.ok=TRUE)
x1=xx=="fahrenheit"
x[x1]=(x[x1]-32)*5/9
x+273.15}
```

With formatting:
```{r}
to_kelvin = function(temp, unit) {
  unit = match.arg(unit, c("celsius", "fahrenheit"), several.ok = TRUE)
  is_f = unit == "fahrenheit"
  
  temp[is_f] = (temp[is_f] - 32) * 5 / 9
  temp + 273.15
}
```

Defensive Programming
---------------------

Test whether inputs satisfy your assumptions.

For instance, if your function should only work on scalars, check the
lengths of the arguments.

Raise an error with `stop()` if the assumptions aren't satisfied.

Useful functions for testing assumptions:

* `is.X()` functions for testing various properties. For instance,
  `is.character()` to test whether an object is a character vector.

* `inherits()` for testing the class of an object.

* `length()`, `dim()` for testing dimensions.



The R Debugger
==============

Debugging code is the process of confirming, step-by-step, that what you belive
the code does is what the code actually does.

The key idea is to check each step of the code.

Two strategies:

* Work forward through the code from the beginning.
* Work backward from the source of an error.

R has built-in functions to help with debugging.

The `browser()` function pauses the running code and starts R's
debugging system.

For example:
```{r}
f = function(n) {
  total = 0

  for (i in 1:n) {
    #browser()
    total = total + i
  }
  
  total
}

f(10)
```

The key debugger commands are:

* `n` to run the next line
* `s` to "step into" a call
* `c` to continue running the code
* `Q` to quit the debugger
* `where` to print call stack
* `help` to print debugger help


Another example:
```{r}
g = function(x, y) (1 + x) * y

f = function(n) {
  total = 0

  for (i in 1:n) {
    #browser()
    total = total + g(i, i)
  }
  
  total
}

f(11)
```


Other Debugging Functions
=========================

The `debug()` function places a call to `browser()` at the beginning
of a function.

Use `debug()` to debug functions that you can't or don't want to edit.

For example:
```{r}
f = function(x, y) {
  x + y
}

debug(f)
```

Use `undebug()` to reverse the effect of `debug()`:
```{r}
undebug(f)
```

The `debugonce()` function places a call to `browser()` at the
beginning of a function for the next call only.

The idea is that you then don't have to call `undebug()`.

Example:
```{r}
debugonce(f)
```

The global option `error` can be used to make R enter the debugger any time an
error occurs.

Set the option to `error = recover`:
```{r}
options(error = recover)
```

Then:
```{r}
f = function(x, y) {
  stop("Hi")
  x + y
}
```

See Chapter 13 of "The Art of R Programming" by N. Matloff for more.
