---
title: "STAT 33A Workbook 1"
date: "Sep 3, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This workbook is due __Sep 3, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. The first lecture video this
week will teach you more about how the notebook works.

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.


Packages & R Notebooks
======================

Watch the "Packages & Notebooks" lecture video.

## Exercise 1

_You don't need to write anything in the notebook for this exercise._

Install the `rmarkdown` and `tinytex` packages.

Confirm that your setup works by knitting this file to PDF.

If you run into any trouble, ask on Piazza or in office hours as soon as
possible. You'll need to be able to knit to turn in the assignment.

_note: I used my existing MikTex distribution_

Markdown
========

Watch the "Markdown" lecture video.

For the remaining exercises, answer questions with complete sentences, and put
code in code chunks. You can make as many new code chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

## Exercise 2

Write 3-5 sentences introducing yourself. Try out some of the Markdown
formatting: italics, bold, lists, quotes, code chunks, and links.

After you finish, knit the notebook to check that your formatting was applied
correctly.

YOUR ANSWER GOES HERE:

Hi! My name is __Ming Fong__ and I'm a first year at Berkeley. Here are some links to personal websites:

* [GitHub](https://github.com/evilpegasus)
* [LinkedIn](https://www.linkedin.com/in/mingfong/)
* [Personal Website](http://mingyfong.github.io/)

1. One fish
2. Two fish
3. Red fish
4. Blue fish

```{r}
x = c(888, 999, 1234)
x = c(x, 111)
x
```

> I use single character variable names so the compiler runs faster.

## Exercise 3

Use Markdown to make:

* A word that's both bold and italic
* A link that's italic
* A nested list (a list with indented subelements)

Try to figure these out by experimentation before asking or searching online.

YOUR ANSWER GOES HERE:

___This is bold and italic___

_[Personal Website](http://mingyfong.github.io/)_

* One
   * One.one
   * One.two
* Two

Vectors
=======

Watch the "Vectors" lecture video.

## Exercise 4

Another way to create vectors is with the `rep()` function. The `rep()`
function creates a vector by replicating a value or vector of values.

1. The first parameter of `rep()` is the thing to replicate. The second
   parameter, `times`, is the number of times to to replicate. Use `rep()` to
   make a vector with 10 elements, all equal to 78.

2. What happens if you pass a vector as the first argument to `rep()`? Give
   some examples.

3. Skim the help file `?rep`. What happens if you pass a vector as the second
   argument to `rep()`? The help file might seem a bit cryptic, so you'll also
   need to experiment. Give some examples.

YOUR ANSWER GOES HERE:

1. 
```{r}
x = rep(78, 10)
x
```

2. It will repeat the argument vector `times` times.
```{r}
y = c(1, 2, 3)
y
z = rep(y, 5)
z
```

3. When the first argument and the `times` argument have the same length, the new vector will repeat the first element `times[1]` times, the second `times[2]` times, and the nth element `times[n]` times.
```{r, eval = FALSE}
?rep
```
```{r}
x = c(1, 2, 3)
x
y = rep(c(1, 2, 3), x)
y
```

## Exercise 5

Yet another way to create vectors is with the `seq()` function. The `seq()`
function creates a vector that contains a sequence of numbers.

Skim the help file `?seq`. Give some examples of creating vectors with the
`seq()` function.

YOUR ANSWER GOES HERE:

```{r, eval = FALSE}
?seq
```
```{r}
x = seq(2, 10, 2)
x
y = seq(2, 100, 2)
y
```

Subsets of Vectors
==================

Watch the "Subsets of Vectors" lecture video.

## Exercise 6

Try out each of the following:

* Getting a single element of a vector
* Reassigning an element of a vector
* Getting multiple elements of a vector
* Getting multiple elements of a vector using negative indices

YOUR ANSWER GOES HERE:

```{r}
x = seq(2, 20, 2)
x
x[5]
x[c(1, 2, 3)]
x[c(-1,-2,-3)]
```

## Exercise 7

Consider this code:
```{r}
x = seq(1, 5) * 10
x[c(3, 4, 5)][1]
```

Try running the code to see what the result is. Then explain what's happening
in the second line (it may be useful to explain what `x[c(3, 4, 5)]` does
first).

YOUR ANSWER GOES HERE:

1. `x = seq(1, 5) * 10` creates a sequence of 1 to 5 and multiplies each value by 10: 10, 20, 30...
2. `x[c(3, 4, 5)]` grabs the third, fourth, and fifth element from x: 30, 40, 50.
3. `x[c(3, 4, 5)][1]` the extra `[1]` grabs the first element after the above step: 30.

## Exercise 8

The `sort()` function sorts the elements of a vector. For instance:
```{r}
x = c(4, 5, 1)
sort(x)
```

Another way to sort vectors is by using the `order()` function. The order
function returns the _indices_ for the sorted values rather than the values
themselves:
```{r}
x = c(4, 5, 1)
order(x)
```

These can be used to sort the vector by subsetting:
```{r}
x[order(x)]
```

The key advantage of `order()` over `sort()` is that it can also be used to
sort one vector based on another, as long as the two vectors have the same
length.

Create two vectors with the same length, and use one to sort the elements of
the other.

YOUR ANSWER GOES HERE:

```{r}
x = seq(2, 10, 2)
x
y = c(1234, 100, 50, 23, 5)
y
# sort x backwards because y is in reverse order
x[order(y)]
```

Submitting Your Work
====================

Congratulations, you made it through the first workbook!

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.
