---
title: "Stat 33B - Lecture Notes 3"
date: September 6, 2020
output: pdf_document
---

File Systems
============

For this section, see the slides for notes.



The R Working Directory
=======================

The **working directory** is the reference point R uses for relative paths.

When navigating files in R, use the working directory to save time and make
code reproducible.

Important functions:

* `getwd()` -- get the working directory
* `setwd()` -- set the working directory
* `list.files()` -- list files in a directory

Try running these in the console first (rather than using `Ctrl` + `Enter` in a
notebook chunk).

Use `getwd()` to get the working directory:
```{r}

```
The output on your computer will probably be different!


Use `setwd()` to set the working directory:
```{r}

```


Use `list.files()` to list files in a directory:
```{r}

```

The output is an empty vector if:

* The path you provided is incorrect.
* The path you provided leads to a file, not a directory.
* There are no files in the directory.

For example, if we make a deliberate typo:
```{r}

```


## R Markdown Files

RStudio tracks the working directory for the console and each R Markdown file
separately.

So:

* Running `getwd()` from the notebook with `Ctrl + Enter` displays the
  NOTEBOOK'S working directory.
* Typing `getwd()` in the "Console" window and pressing `Enter` displays the
  CONSOLE's working directory.

In the notebook, if you use `setwd()` it only lasts for __that chunk__ and is
then reset:
```{r}

```

So in subsequent chunks it looks like you didn't call `setwd()`:
```{r}

```

Why does RStudio do this? It is a bad practice to include `setwd()` in your
notebooks, because people you share the notebook with, like your colleagues,
instructor, or employer, might not have the same directories on their computer
as the ones you have on your computer. The next section has more details about
this.

By default, RStudio does the right thing and sets the notebook's working
directory to the place where the notebook is saved. Then you can use relative
paths (see below) to load and save files from the
notebook.

If you really want to set the working directory in a notebook, it is possible
to override RStudio. See <https://yihui.org/knitr/options/> for details.


## Editing Paths

R also has functions to make it easier to edit/create paths:

* `normalizePath()` -- convert relative path to absolute path
* `file.path()` -- combine parts of a path
* `dirname()` -- get all except last component of path
* `basename()` -- get last component of path

You can use `normalizePath()` to inspect the path shortcuts:
```{r}

```

The path `~` is your **home directory**:
```{r}

```
Your home directory is probably different!


The `file.path()` function combines parts of a path:
```{r}

```

The `dirname()` and `basename()` functions get parts of a path:
```{r}

```


## Reproducible Analyses

Plan ahead so that other people can run your code and reproduce your results.

Good habits:

* Putting your notebook(s) and data in the project directory.
* Using paths relative to the project directory.

Bad habits:

* Calling `setwd()` in R notebooks and scripts.
* Using absolute paths.

It's okay to use `setwd()` in the *R console* to set the working
directory to your project directory.




















Data Frames
===========

The first step of an analysis is to load a data set.

Many different file formats exist for storing data. The broadly fall into two
categories:

1. **Plain-text**, which means the format is human-readable. You can open and
   edit these in any text editor.

2. **Binary**, which means the format is not human-readable. You need specific
   software to open and edit these. Compared to plain-text, most binary formats
   are faster to read and write, and use less space.


R provides a binary data format called **RDS** (R data, serialized). The
extension on an RDS file is usually `.rds`.

Any R object can be stored an RDS file.

Use `saveRDS()` to save an object to an RDS file:
```{r}

```
This is a good way to save your work after a long computation.


Use `readRDS()` to load an object from an RDS file:
```{r}

```


## Data Frames

In statistics, we frequently work with 2-dimensional tables of data.

For a tabular data set, typically:

  * Each row corresponds to a single case or subject. These are called
    **observations**.
  * Each column corresponds to something the data measures. These are called
    **features**, **covariates**, or variables.

A "variable" means something else in R, so I'll avoid using it to refer to
columns.

R's data structure for tabular data is the **data frame**.

Let's load a data frame:
```{r}

```

This data set is available on the bCourse.

The Dogs Data Set is based on:

    https://informationisbeautiful.net/visualizations/
        best-in-show-whats-the-top-data-dog/


## Inspecting Objects

Printing is not the only way to inspect data, and has drawbacks:

1. Slow (especially if you're knitting a notebook)
2. Hard to read if there are lots of columns

R provides functions to inspect objects.

We already saw one of these:
```{r}

```

Use `head()` to print the first 6 rows (or elements):
```{r}

```

Use `tail()` for the last 6:
```{r}

```

Use `dim()` to print the dimensions:
```{r}

```

Alternatively, use `ncol()` and `nrow()`:
```{r}

```

Use `names()` to print the column (or element) names:
```{r}

```

Use `rownames()` to print the row names:
```{r}

```

Use `str()` to print a structural summary:
```{r}

```

Use `summary()` to print a statistical summary:
```{r}

```


## More about Data Frames

R uses data frames to represent tabular data.

A data frame is a list of column vectors. So:

* Elements of a column must all have the same type (like a vector).
* Elements of a row can have different types (like a list).
* Every row must be the same length.

In addition, every column must be the same length.


This idea is reflected in the type of a data frame:
```{r}

```


## Accessing Columns

Recall that lists can have named elements:
```{r}

```

The dollar sign operator `$` extracts a named element from a list.

It's especially useful for getting columns from data frames:
```{r}

```


You can also use `$` to set an element:
```{r}

```


## Deconstructing Data Frames

The `unclass()` function resets the class of an object to match the object's
type.

You can use `unclass()` to inspect the internals of an object.

For example, you can see that a data frame is a list:
```{r}

```

















Factors
=======


R represents categorical data using the class `factor`:
```{r}

```

The categories of a factor are called **levels**.

You can list the levels with the `levels()` function:
```{r}

```


Factors remember all possible levels even if you take a subset:
```{r}

```

This is one way factors are different from strings.

For example:
```{r}

```


You can make a factor forget levels that aren't present with `droplevels()`:
```{r}

```


You can create a factor with the `factor()` function:
```{r}

```


## Counting Things

The `table()` function returns the frequency of each value in a vector.

This is especially useful for factors:
```{r}

```


## Deconstructing Factors

Internally, R represents factors as integer vectors:
```{r}

```
Each integer corresponds to one level of the factor.

This representation uses less memory than repeating each level's
name.























File Formats
============

Recall there are two kinds of file formats: plaintext and binary.

The RDS format is a binary format.


## Plaintext Formats for Tabular Data

Several plaintext file formats are designed just for tabular data:

* Delimited files
    + Comma-separated value (CSV) files
    + Tab-separated value (TSV) files
* Fixed-width files


For example, suppose you download the Significant Volcanic Eruption Database
from:

    https://www.ngdc.noaa.gov/nndc/struts/form?t=102557&s=50&d=50


This file is also on the bCourse.


The website says the file is tab-delimited, so use `read.delim()`:
```{r}

```

Many things can go wrong when you read tabular data from a plaintext file:

* Extra lines in the file
* No header in the file
* Incorrect column classes

These can generally be fixed by setting parameters in the read function.

For example, to skip the first 10 lines and disable the header:
```{r}

```

You can read more about the parameters in `?read.table`.


The most common plaintext data format is CSV.

Use `write.csv()` to write CSV files:
```{r}

```

For other formats:

* `read.csv()` -- read CSV files
* `read.table()` -- read delimited files in general
* `read.fwf()` -- read fixed-width files


## Binary Formats for Tabular Data

There are also a few binary formats for tabular data:

* Excel spreadsheets
* Feather

R doesn't provide built-in support for these, but check CRAN for packages.

For instance:

* `readxl` -- a package with functions for reading Excel spreadsheets
* `arrow` -- a package with functions for reading Feather files


## Non-tabular Data

Many packages are available for non-tabular file formats.

For example:

* `jsonlite` -- read JavaScript Object Notation (JSON) files
* `xml2` -- read Extensible Markup Language (XML) files
* `rvest` -- read Hypertext Markup Language (HTML) files

The built-in `readLines()` function can read lines from any plaintext file.

For example:
```{r}

```

Think of `readLines()` as a last resort.
