---
title: "STAT 33B Homework 1"
date: "Sep 10, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This homework is due __Sep 10, 2020__ by 11:59pm PT.

Homeworks are graded for correctness. 

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.


## Exercise 1

Note: In R Markdown formatting, a dollar sign `$` marks the beginning of a
(LaTeX) mathematical expression. For example, to format a linear equation
nicely, you can write $ax + b = 0$. You don't need to know LaTeX for this
class, so if you don't understand something written inside `$`, knit the PDF
and read that instead.

The `curve()` function plots a curve based on an expression. The basic syntax
is `curve(expr, from, to)` where `from` and `to` are the limits of the x-axis.

For example, to plot $x^2$ between 0 and 3:
```{r}
curve(x^2, 0, 3)
```

The `curve()` function makes it possible to use R as a graphing calculator.

For each expression below:

1. Plot the expression so that all points where it is 0 are visible. Experiment
   to find appropriate limits for the x-axis.

2. At which point(s) is the expression zero? Estimate (within 1 unit) based on
   your plot rather than computing these mathematically.

Here are the expressions:

1. $5x - 1$
2. $3x^2 - 2x - 8$
3. $\sqrt{3x} - 4$

### YOUR ANSWER GOES HERE:

1. Zero at approximately $x = 0.2$
```{r}
curve(5*x - 1, 0, 2)
```

2. Zeros at approximately $x = -1.5, 2$
```{r}
curve(3*x^2 - 2*x - 8, -5, 5)
```

3. Zero at approximately $x = 0$
```{r}
curve(sqrt(3*x), 0, 5)
```

## Exercise 2

A **discrete probability distribution** is a table of mutually exclusive
outcomes and their associated probabilities.

For example, the **discrete uniform distribution** assigns equal probability to
each outcome. So the discrete uniform distribution on the integers 1-4 is:

Outcome | Probability
------- | -----------
1       | 0.25
2       | 0.25
3       | 0.25
4       | 0.25

A fair coin toss is another instance of a discrete uniform distribution, where
each of the two outcomes has probability 0.5.

You can plot the distribution above with the code:
```{r}
outcome = seq(1, 4)
prob = rep(0.25, 4)
plot(outcome, prob)
```

Because the probabilities are all equal, a plot of a uniform distribution is
always flat, regardless of the number of outcomes.

The idea of a uniform distribution can also be extended to continuous values.
For example, suppose we want to select a decimal value between 0 and 4
(inclusive), with equal probability for any value in the interval. This is a
**countinuous uniform distribution**.

As with any uniform distribution, a plot of a continuous uniform distribution
is flat. However, instead of discrete points, the distribution is a curve along
an interval.

You can plot the continuous uniform distribution from 0 to 4 with the code:
```{r}
curve(dunif(x, 0, 4), -1, 5, n = 1000)
```

A plot of a continuous distribution is called a **density plot**.

In a density plot, the y-axis no longer represents probability. Instead, the
probability of points in any subinterval is the total area between the curve
and the line $y = 0$ over the subinterval. For instance, in the distribution
above, the probability of the outcome being a point between 1 and 2 (inclusive)
is 0.25.

R provides functions to compute the density of various distributions. The code
above uses the `dunif()` function, the **d**ensity plot function for the
**unif**orm distribution. The function has parameters to control the interval
for the distribution.

R also provides functions to sample values randomly from various distributions.
For example, the `runif()` function samples **r**andomly from a continuous
**unif**orm distribution. You can sample 10 values from the continuous uniform
distribution from -1 to 1 with the code:
```{r}
runif(10, -1, 1)
```

Since these values are randomly sampled, they will be different each time you
run the code.

The **Gaussian distribution** or normal distribution is another continuous
probability distribution. The distribution plays an important role in
mathematics, statistics, and the natural sciences. The density plot for this
distribution is often described as a "bell-shaped curve". The Gaussian
distribution is typically parameterized by its mean (center) and standard
deviation (how spread out it is), rather than the interval it covers.

R provides functions `dnorm()` and `rnorm()` for the Gaussian distribution that
are analogous to the `dunif()` and `runif()` functions described above.

1. Use the `curve()` function to make a density plot of the Gaussian
   distribution with mean 10 and standard deviation 5. Set the x-axis to span
   from -10 to 30.

2. Skim the documentation for `rnorm()`. Then create a variable called `samp`
   that contains 10 random values from a normal distribution with mean 10 and
   standard deviation 5.

3. Compute the mean and standard deviation of `samp`. Comment briefly on how
   much these differ from the true mean 10 and true standard deviation 5.

   What happens if you run your code a second time? Are the sample mean and
   standard deviation the same? Are the sampled points the same?


### YOUR ANSWER GOES HERE:

1. 
```{r}
curve(dnorm(x, 10, 5), -10, 30)
```

2. 
```{r}
mu = 10
sigma = 5
samp = rnorm(10, mu, sigma)
samp
```

3. Running the code many times will change the random sample and also change the sample mean and sd. The difference is computed below.
```{r}
mu2 = mean(samp)
mu2
# The mean of the sample differs from the orignal mean by
mu2 - mu

sigma2 = sd(samp)
sigma2
# The standard deviation of the sample differs from the orignal standard deviation by
sigma2 - sigma
```

## Exercise 3

R's functions for random sampling use a pseudo-random number generator (PRNG).
The numbers produced by a PRNG are not truly random, but satisfy conditions
that make them close enough to random for most scientific computing tasks.

An advantage of PRNGs is that they are deterministic. Given the same initial
parameter, called a **seed**, a PRNG will generate the same sequence of
"random" numbers every time. When you start R, the seed is automatically
assigned value from your computer that varies (e.g., the system time in
milliseconds). This ensures that you get a different sequence of random numbers
in each R session.

You can the PRNG seed with the `set.seed()` function. Setting the seed makes
your results reproducible. That is, other people can run your code and get the
same results even if the code generates random numbers.

Generally, you should only set the seed once at the beginning of an R script or
notebook, because it affects all subsequent calls that use the PRNG.

1. Set the seed to 93.

2. Create a variable called `samp` that contains 1000 random values from a
   normal distribution with mean 10 and standard deviation 5.

3. Compute the mean and standard deviation of `samp`.

   What happens if you run your code a second time? Are the sample mean and
   standard deviation the same? Are the sampled points the same? Explain how
   step 1 affects this.

4. Compute the number of values in `samp` that are greater than 8. _Hint: there
   is a fast, simple way to count `TRUE` values using implicit coercion._

### YOUR ANSWER GOES HERE:

1. 
```{r}
set.seed(93)
```

2. 
```{r}
samp = rnorm(1000, 10, 5)
```

3. If you set the seed and run the code again, the sampled points are always the same. `set.seed()` ensures the pseudo-random numbers used in `rnorm()` are the same each time.
```{r}
mean(samp)
sd(samp)
```

4. 
```{r}
sum(samp > 8)
```

## Exercise 4

A **Monte Carlo algorithm** is an algorithm that uses random number generation
to approximate a non-random result.

One application of Monte Carlo algorithms is to approximate the area of a
shape. In this exercise, you'll use a Monte Carlo algorithm to approximate the
area of a circle, and thereby approximate pi.

Consider the circle centered at 0 with radius 1. Points on the circle satisfy
$x^2 + y^2 = 1$. This is the "target shape" for which you'll approximate the
area. In practice, the target shape is usually one whose area is difficult to
compute by hand. 

In order to approximate the area of the target shape, start by finding a
square or rectangle that completely encloses the target shape. For instance, we
can use the square with corners (-1, -1) and (1, 1) to enclose the circle.

Next, randomly sample x coordinates and y coordinates uniformly and
independently over the square.

Now the proportion of sampled points inside the circle corresponds to the ratio
between the area of the circle and the area of the square. Since you know the
area of the square, you can use this relationship to compute the area of the
circle.

1. Use R and the described algorithm to approximate the value of pi. Use a
   variable called `n` to control the sample size.

   Do not use if-statements or loops in your solution; they are not necessary.
   Try to take advantage of vectorization and implicit coercion. A typical
   solution will be 2-6 lines.

2. Test your code for different values of `n` from 10 to 1,000,000. How does
   changing `n` affect the accuracy of the approximation? How does it affect
   the time it takes to compute the approximation?

   _Hint 1: R supports exponent notation, so you can write 1,000,000 in R as
   `1e6`._

   _Hint 2: The `microbenchmark` package provides a simple function called
   `microbenchmark()` for timing expressions. You can use it to time an entire
   block of code by surrounding the code with curly braces `{ }`._

### YOUR ANSWER GOES HERE:
