---
title: "STAT 33B Workbook 2"
date: "Sep 10, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This workbook is due __Sep 10, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.


File Systems
============

Watch the "File Systems" lecture video.


## Exercise 1

For each of the following paths, say whether the path is absolute or relative,
and explain how you can tell.

1. `"./documents"`
2. `"/Users/Jun/doggos_to_pet.md"`
3. `"TODO.md"`

YOUR ANSWER GOES HERE:

1. Relative. The path starts with a `.` which searches for `"/documents"` from the working directory

2. Absolute. The path starts at the root `/`.

3. Relative. The path searches for `"TODO.md"` from the current working directory.

The R Working Directory
=======================

Watch the "The R Working Directory" lecture video.


## Exercise 2

1. What's the root directory called on your computer?
2. What's the absolute path to the home directory on your computer?
3. Use R to count the total number of files in your home directory. Your code
   should return the result as a number.

YOUR ANSWER GOES HERE:

1. `"C:/"`

2. `""C:\\Users\\mingf\\Documents""`
```{r}
normalizePath("~")
```

3. There are 14 files in my home directory. See below:
```{r}
length(list.files("~"))
```

Data Frames
===========

Watch the "Data Frames" lecture video.

## Exercise 3

1. Load the dogs data set `dogs.rds` into R (this one has more than 10 rows).
2. What's the mean weight of the dogs? You can use the `na.rm` parameter in the
   `mean()` function to make the function ignore missing values.
3. The `which.min()` function returns the index of the minimum element of a
   vector. Which breed of dog has the shortest height?
4. Which breed of dog has the longest lifespan?

YOUR ANSWER GOES HERE:

1. 

2. 

3. 

Factors
=======

Watch the "Factors" lecture video.


No exercises for this video. Get up, stretch, and take a break! :)


File Formats
============

Watch the "File Formats" lecture video.


## Exercise 5

1. Load the volcano data set into R.
2. What are the column names? Use R to get these rather than typing them out
   yourself.
3. How many volcano eruptions are recorded in the data set?
4. What are the classes/types of the columns? _Hint: an earlier lecture
   mentioned a function that summarizes of this information._
5. Are there any columns that contain categorical data? Are these columns
   factors? If not, what are their classes?

YOUR ANSWER GOES HERE:
