---
title: "Stat 33B - Lecture Notes 13"
date: November 15, 2020
output: pdf_document
---


Tidy Data
=========

Many Tidyverse functions require data in _tidy_ form:

1. Each observation must have its own row.
2. Each feature must have its own column.
3. Each value must have its own cell.

Tidy data sets are convenient for data analysis in general.


The tidyr package provides functions to reshape data to and from tidy form.

An example of tidy data (Tuberculosis Data Set):
```{r}
# install.packages("tidyr")
library(tidyr)

table1
```

The `country` and `year` columns identify the observation.

The `cases` and `population` columns are measurements (features).


Determine the identifiers and the measurements are BEFORE reshaping!





Columns into Rows
=================

Tidy data rule 1 says each observation must have its own row.

Here's a table that breaks rule 1:
```{r}
library(tidyr)

table4a
```
All of the numbers measure the same thing: cases.


To make the data tidy, we must rotate the `1999` and `2000` column names into
rows, one for each value in the columns. New columns are `year` and `cases`.


Less columns (generally), more rows. The data set becomes longer.


Use `pivot_longer` to rotate columns into rows.

Need to specify:

* Columns to rotate as `cols`.
* Name(s) of new identifier column(s) as `names_to`.
* Name(s) of new measuerment column(s) as `values_to`.

Here's the code:
```{r}

```


You also can do this without `tidyr`:

1. Subset columns to separate `1999` and `2000` into two data frames.
2. Add a `year` column to each.
3. Rename the `1999` and `2000` columns to `cases`.
4. Stack the two data frames with `rbind`.

```{r}

```





Rows into Columns
=================

Tidy data rule 2 says each feature must have its own column.


Let's look at a table that breaks rule 2:
```{r}
library(tidyr)

table2
```
Here the `count` column contains two different features: cases and population.


To make the data tidy, we must rotate the `count` values into columns, one for
each `type` value. New columns are `cases` and `population`.


Less rows, more columns. The data set becomes wider.


Use `pivot_wider` to rotate rows into columns.

Need to specify:

* Column names to rotate as `names_from`.
* Measurements to rotate as `values_from`.

Here's the code:
```{r}

```


You can also do this without `tidyr`:

1. Subset rows to separate `cases` and `population` values.
2. Remove the `type` column from each.
3. Rename the `count` column to `cases` and `population`.
4. Merge the two subsets by matching `country` and `year`.

```{r}

```

We'll see a better way to do step 4 later.


Run `vignette("pivot")` for more examples of how to use tidyr.





String Processing
=================

Tidy data rule 3 says each value must have its own cell.

Here's a table that breaks rule 3:
```{r}
library(tidyr)

table3
```
Cells in the `rate` column contain two values: cases and population.

The values in the `rate` column are strings:
```{r}

```

So we need a way to manipulate strings.

Some things we might want to do with strings are:

* Detect one string within another
* Split a string into parts
* Extract part of a string
* Replace one string within another

There are base R functions for working with strings, but they have an
inconsistent interface.

The stringr package (part of the Tidyverse) also provides functions for
searching and transforming strings.

We'll use stringr:
```{r}
# install.packages("stringr")
library(stringr)
```

The typical syntax of stringr functions is:
```
str_FUNCTION(string, pattern, ...)
```

For example, the `str_detect` function detects whether the pattern appears
within the string:
```{r}

```

The stringr functions are vectorized:
```{r}

```

The stringr functions use a special language called _regular expressions_ or
_regex_. In regex, some characters have special meanings.

You can disable regular expressions with the `fixed` function:
```{r}

```
We'll learn more about regex in a different lecture.


The `str_replace` function replaces the pattern the first time it appears in
the string:
```{r}

```


The `str_replace_all` function replaces the pattern every time it appears in
the string:
```{r}

```


The `str_split_fixed` function splits a string into a fixed number of pieces,
at some separating character:
```{r}

```

We can use `str_split_fixed` to tidy the data from the table:
```{r}

```





Regular Expressions
===================

Patterns in stringr use a language called _regular expressions_ or _regex_.

A regular expression can describe a complicated pattern in just a few
characters, because some characters, called _metacharacters_, have special
meanings.

Letters and numbers are NEVER metacharacters. They are always literal.

Metacharacter | Matches
------------- | -------
`.`           | any character
`^`           | beginning of string
`$`           | end of string
`[ab]`        | `'a'` or `'b'`
`[^a]`        | any character except `'a'`
`*`           | previous character appears 0 or more times
`+`           | previous character appears 1 or more times
`?`           | previous character appears 0 or 1 times
`()`          | make a group
`(ab|c)`      | match `'ab'` or `'c'`

More metacharacters are listed on the stringr cheatsheet, or in:
```{r}
?regex
```


We can test these with `str_detect`.

For example, the wildcard `.`:
```{r}

```


Square brackets create a _character class_, a set of characters where any one
will match:
```{r}

```


Parentheses create a _group_:
```{r}

```
Notice the parentheses are not matched literally.


The `*`, `+`, and `?` are _quantifiers_ that repeat a character or group:
```{r}

```
They are _greedy_, meaning they match as many characters as possible.


The `str_match` and `str_match_all` functions extract matching groups:
```{r}

```
