---
title: "Stat 33B - Lecture Notes 10"
date: October 25, 2020
output: pdf_document
---


Environments
============

Motivation: How does scope in R work (technical details)?


An _environment_ is a collection of named R objects.

Each environment has two parts:

1. A _frame_ that associates names with R objects (a hash map)
2. A _parent environment_


Use `new.env` to create a new environment:
```{r}

```

Use `ls` or `names` to list the names in an environment:
```{r}

```

Use `exists` to check for a specific name:
```{r}

```

Use `[[` or `$` to get or set elements of an environment:
```{r}

```


By default, the `ls` function ignores names that start with `.`

Use `all.names = TRUE` to make the function print these names:
```{r}

```


Environments are _reference objects_.

So they **do not** follow the copy-on-write rule:
```{r}

```



Variable Lookup, Part 2
=======================

R uses environments to store variables.

At run-time, **each call** to a function creates a new environment:

* The frame contains local variables.
* The parent environment is the environment where the function was
  **defined** (lexical scoping).

To look up a variable, R checks the local environment, then the parent, then
the parent's parent, and so on.


Use `parent.env` to get the parent environment:
```{r}

```

The _global environment_ stores variables you define in the console (not in
functions).

Use `globalenv` to get the global environment:
```{r}

```

Most functions for manipulating environments default to the local environment.

Use `environment` to get the local environment:
```{r}

```

Typing a variable name gets an element in the local environment:
```{r}

```

The `get` function can also get an element:
```{r}

```
Use `inherits = FALSE` to ignore the parent environment.


The `<-` and `=` operators set an element in the local environment:
```{r}

```

The `assign` function can also set an element:
```{r}

```



The Search Path
===============

R uses environments to store variables in packages, too.

When R loads a package, it creates a new environment:

* The frame contains the package's local variables.
* The parent environment is the environment of the previous package loaded.
* This new environment becomes the parent of the global environment.

R always loads several built-in packages at startup.


So the global environment is not the top-level environment:
```{r}

```

The _base environment_, or `package:base`, corresponds to the built-in R
package `base`.

The parent of the base environment is a special environment called the _empty
environment_, which contains no variables and has no parent.

Use `baseenv` and `emptyenv` to access these:
```{r}

```

The history of loaded packages is called the _search path_.

Use `search` to see the search path:
```{r}

```
Package environments (and the global environment) are listed from child to
parent.




The Colon Operators
===================

The double-colon operator `::` gets a variable in a package.

Two common uses:

1. Disambiguate which package you mean, when packages have variables with the
   same names.
2. Get a variable from a package without loading the package.


For example:
```{r}

```

The related triple-colon operator `:::` gets a _private_ variable in a package.

Generally these are private for a reason!

Only use `:::` if you're sure you know what you're doing.



Closures
========

A _closure_ is a function together with an enclosing environment.

Every R function is a closure.

The enclosing environment is the environment where the function was
**defined**.

Call `environment()` on a function to get its enclosing environment:
```{r}

```

The enclosing environment persists between calls.

You can use this to make a function that "remembers" previous calls.


Use the superassignment operator `<<-` to assign a variable to the local
environment's parent.

Inside a function, the local environment's parent is the enclosing environment.

So `<<-` is especially useful inside functions:
```{r}

```

As an example, let's make a function that returns the number of times it's been
called.

Here's the code:
```{r}

```

Modifying the enclosing environment is a _side effect_.

Functions with side effects make code harder to understand and predict.

Use side effects sparingly. Most functions should not have side effects.

If you do need side effects, try to isolate them.


Creating Closures
-----------------

The example function has side effects on the global environment.

This extremely bad design!

The function might overwrite the user's variables:
```{r}

```

Or the user might overwrite the function's variables:
```{r}

```

It's better to create a isolated enclosing environment for the function.

There are two different idioms for doing this:

1. Define the function inside of another function (a _factory function_)
2. Define the function inside of a call to `local`


Here's skeleton code for the first approach:
```{r}
make_f = function() {
  # Put variables that `f` remembers between calls here:

  # This is your closure `f`:
  function() {
    # ...
  }
}

f = make_f()
# Now you can use f() as you would any other function.
```

We can use this approach for the example function:
```{r}

```
