---
title: "STAT 33B Workbook 9"
date: "Oct 29, 2020"
author: "YOUR NAME (YOUR SID)"
output: pdf_document
---

This workbook is due __Oct 29, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

If you have any last-minute trouble knitting, **DON'T PANIC**. Submit your Rmd
file on time and follow up in office hours or on Piazza to sort out the PDF.


Environments
============

Watch the "Environments" lecture video.

No exercises for this section.



Variable Lookup, Part 2
=======================

Watch the "Variable Lookup, Part 2" lecture video.

No exercises for this section.



The Search Path
===============

Watch the "The Search Path" lecture video.

## Exercise 1

Create a function called `locate` that finds and returns the first
environment (in a chain of environments) that contains a given variable
name. Your function should have a parameter `name` for the quoted variable
name and a parameter `env` for the initial environment to search.

If the variable is not present in any of the environments in the chain, your
function should return the empty environment.

_Hint 1: Use `exists` to check whether the variable exists in `env` (and not
its ancestors). If it does, return `env`. If it does not, set `env` to be
its own parent and repeat this process._

_Hint 2: You can use `identical` to check if two environments are equal._

**YOUR ANSWER GOES HERE:**



## Exercise 2

This code produces an environment `e` with several ancestors:
```{r}
e = new.env()
e$c = 42
e = new.env(parent = e)
e$a = "33a"
e$b = "33b"
e$c = "33ab"
e = new.env(parent = e)
e$x = 8
```

Test your `locate` function on `e` by:

1. Locating `"c"`. Your result `result` should have `result$c` equal to
   `"33ab"`.

2. Locating `"zzz"`. Your result `result` should be the empty environment.

3. Locating `"e"`. Your result `result` should be the global environment.

4. Locating `"show"`. Which built-in package provides this function?

**YOUR ANSWER GOES HERE:**

### Part 1

### Part 2

### Part 3

### Part 4



The Colon Operators
===================

Watch the "The Colon Operators" lecture video.

No exercises for this section.



Closures
========

Watch the "Closures" lecture video.

## Exercise 3

Recall the `find_fib2` function (from week 8) for computing Fibonacci numbers:
```{r}
find_fib2 = function(n, fib = c(1, 1)) {
  len = length(fib)
  if (n <= len)
    return (fib[n])
  
  fib = c(fib, fib[len - 1] + fib[len])
  Recall(n, fib)
}
```

The key to computing Fibonacci numbers efficiently is to keep a record of the
numbers that have already been computed. The `find_fib2` function does this by
passing the record of computed numbers on through the `fib` parameter.

An alternative to passing the record through a parameter is to store the record
in the function's enclosing environment. Write a function `find_fib3` that does
this.

Your function should still have a parameter `n` and return the `n`-th Fibonacci
number. Your function should **NOT** have a parameter `fib`.

Test your function by computing `fib(40)`. If your function is working
correctly, it should be able to compute this number in less than 5 seconds, and
the number should be `102334155`.

_Hint 1: Create a factory function `make_find_fib3` to provide the enclosing
environment. The factory function should not have any parameters, and should
return your `find_fib3` function._

_Note: Using the enclosing environment to store values that have been computed
is called "memoization". Memoization is a useful strategy for improving
efficiency in many programming problems._

**YOUR ANSWER GOES HERE:**



## Exercise 4

1. Enclosing environments persist between calls. Explain how you think this
   will affect the speed and memory usage of `find_fib3` compared to
   `find_fib2`. What are the advantages and disadvantages of the two different
   functions?

   _Hint: How long does it take to compute `find_fib3(40)` the first time? The
   second time? The third time?_

2. Use the microbenchmark package to benchmark `find_fib2` and `find_fib3` for
   `n` equal to 30 and `n` equal to 40. Which function is faster?

**YOUR ANSWER GOES HERE:**

### Part 1

### Part 2
