---
title: "STAT 33B Homework 2"
date: "Sep 24, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This homework is due __Sep 24, 2020__ by 11:59pm PT.

Homeworks are graded for correctness. 

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.

Exercise 1
----------

For this assignment, you'll use the Datasaurus Dozen data set, which is
available on the bCourse (`DatasaurusDozen.tsv`).

Load the Datasaurus Dozen data set and assign it to a variable named `dsaur`.

**YOUR ANSWER GOES HERE:**

```{r}
dsaur = read.delim("data/DatasaurusDozen.tsv", header = TRUE)
```

Exercise 2
----------

Now that you've loaded the data set, print out summary information, including:

+ Number of columns
+ Number of rows
+ Classes of the columns
+ Levels in the `dataset` column
* The range of the `x` column
* The range of the `y` column
* Number of missing values in each column

**YOUR ANSWER GOES HERE:**

```{r}
ncol(dsaur)
nrow(dsaur)
unlist(lapply(dsaur, class))
levels(factor(dsaur$dataset))
range(dsaur$x)
range(dsaur$y)
sum(is.na(dsaur$dataset))
sum(is.na(dsaur$x))
sum(is.na(dsaur$y))
```

Exercise 3
----------

The Datasaurus Dozen is actually a collection of 12 data sets stacked together.
The `dataset` column indicates which data set each row comes from.

1. Use subsetting to extract only the rows in the `dino` data set. Assign those
   rows to the `dino` variable.

2. Compute the mean and standard deviation for the `x` and `y` columns in the
   `dino` data set.

3. Repeat part 3.1 and 3.2 for the `star` dataset.

   Based on the statistics, are the two data sets similar?

**YOUR ANSWER GOES HERE:**

```{r}
dino = subset(dsaur, dataset == "dino")
mean(dino$x)
sd(dino$x)
mean(dino$y)
sd(dino$y)

star = subset(dsaur, dataset == "star")
mean(star$x)
sd(star$x)
mean(star$y)
sd(star$y)
```
Both the `dino` and `star` datasets have very similar means and SDs. However this could be missleading.

Exercise 4
----------

_Note: Exercise 4-5 use ggplot2, which will be covered in the week 5 lectures._

1. Use `ggplot2` to make a scatter plot of `x` versus `y` for the `dino` data
   set. Make sure your plot includes a title.

2. Repeat for the `star` data set.

   Based on these plots, are the two data sets similar?

**YOUR ANSWER GOES HERE:**

```{r}
library(ggplot2)
ggplot(dino, aes(x = x, y = y)) + geom_point() + labs(title = "dino")
ggplot(star, aes(x = x, y = y)) + geom_point() + labs(title = "star")
```
The plots of the two datasets are not similar. `dino` plots a dinosaur image while `star` plots a five-pointed star.

Exercise 5
----------

A "faceted" plot is one that shows several subplots side-by-side, to aid
comparison between them. Each subplot is called a "facet".

You can create a faceted plot with ggplot2 by using the facet layer. For
instance, the `facet_wrap()` function creates a line of facets based on a
single categorical variable. The facet layer should be added to a plot _after_
the geometry layers.

1.  Read the documentation for `facet_wrap()`, then create a faceted scatter
    plot that shows each dataset from the Datasaurus Dozen in a separate facet.
    Use `geom_smooth` with `method = "lm"` to add a linear regression line to
    each facet.

    _Hint: Unlike other ggplot2 functions, variable names in facet functions
    need to be enclosed in a call to the `vars()` function. So to write the
    column `dataset`, you would write `vars(dataset)`. See the `facet_wrap()`
    documentation for more details._

2. Is there any pattern to the regression lines across the different data sets?

**YOUR ANSWER GOES HERE:**

1. 
```{r}
plot = ggplot(dsaur, aes(x = x, y = y)) +
   geom_point() + labs(title = dsaur$dataset) +
   geom_smooth(method = "lm")

plot + facet_wrap(vars(dataset))
```

2. The regression lines of all the plots are very similar if not identical. All the lines have a slight negative slope. However, the actual points of each dataset are completely different from each other. The very similar statistical summaries masks this.