---
title: "Stat 33B - Lecture Notes 1"
date: August 23, 2020
output: pdf_document
---

R Basics
========

R has a Read-Eval-Print Loop (REPL):

1. Type an expression at the R prompt and hit the enter key.
2. R reads the expression.
3. R evaluates the expression to compute a result.
4. R prints the result in the console.
5. R loops back to waiting for you to enter an expression.

This is similar to Python, Julia, Lisp, etc.

R has many built-in functions for doing math and stats.


# Getting Help

R has built-in documentation.

You can use the `?` command to get help with a specific function:
```{r}

```

You can use the `??` command to search the documentation:
```{r}

```

Strings use single or double quotes (there's no difference).


The help commands work with strings or unquoted names.
```{r}

```


The `sessionInfo()` function prints info about your R session:
```{r}

```


## Order of Operations

The order of operations in R is similar to math and most programming languages.

To see the entire order:
```{r}
?Syntax
```

## Functions & Calls

Recall that:

* __Parameters__ are the inputs a function accepts.
* __Arguments__ are the values assigned to parameters in a call.

You can set arguments by position or by name:
```{r}

```


## Copy-on-write, Part 1

In R, most objects are __copy-on-write__.

That is, if we assign `x` to `y`:
```{r}

```

And then change `x`:
```{r}

```

Then `y` remains unchanged:
```{r}

```

Originally, `x` and `y` referred to the same value in memory.

When we changed `x` (a "write"), R automatically copied the original value so
that `y` remained the same.



Packages & Notebooks
====================

A __package__ is collection of functions and/or data for use in R.

The Comprehensive R Archive Network (__CRAN__) stores most user-contributed
packages.


You can install packages from CRAN with `install.packages()`.

For example:
```{r}

```

A package only needs to be installed once.

For maintaining your packages, there are also the functions:

* `remove.packages()` to remove a package
* `update.packages()` to update ALL packages
* `installed.packages()` to list installed packages


## Loading Packages

The `library()` function loads an installed package:
```{r}

```

Only load the packages you actually need.

You'll have to reload the packages each time you restart R.


## Notebooks

Two typical ways to save R code:

* R script (.R file)
* R notebook (.Rmd file)

R scripts are simpler:

* No extra packages required
* Ideal for developing software

R notebooks are richer:

* Can store formatted text and code
* Can be converted to HTML, DOCX, and PDF
* Ideal for data analyses and presentations

R notebooks require the `rmarkdown` package:
```{r}

```

Generating a report from an R notebook is called __knitting__.


## TinyTeX

If you want to knit PDFs from R notebooks, you also need LaTeX.

LaTeX is programming language for typesetting books.

The `tinytex` package aims to make installing LaTeX easy.

First, install `tinytex`:
```{r}
install.packages("tinytex")
```

Second, tell `tinytex` to install LaTeX:
```{r}
library(tinytex)
install_tinytex()
```
This may take a while, and you may need administrator permissions.

Finally, restart R and try knitting an R notebook.

Remember that the output type must be `pdf_document`.



Vectors
=======

R has no concept of scalars or arrays.

R's atomic data type is the __vector__, an ordered container for 0 or more
elements.

Vector elements must all have the same data type.

The `c()` function combines vectors:
```{r}

```

## Vectorization

A __vectorized__ function is one that is applied element-by-element when passed
a vector argument.

Many R functions are vectorized:
```{r}

```

Vectorization is the fastest kind of iteration in R.



## Indexing

In R, indexes start at 1.

Use the square bracket `[` to access elements of a vector:
```{r}

```

You can use a vector as an index:
```{r}

```


## Copy-on-write, Part 2

The copy-on-write rule applies to vectors.

For example:
```{r}

```

This is different from languages like C and Python.
