---
title: "Stat 33B - Lecture Notes 2"
date: August 30, 2020
output: pdf_document
---

Data Types
==========

In statistics, we categorize data into different types:

* Continuous (real numbers)
* Discrete (integers, or finite number of values)
* Logical (1 and 0s, T and Fs)
* Nominal (categorical values with no ordering)
* Ordinal (categorical values with ordering)
* Graph (network data)
* Textual (books, websites, etc)

R also categorizes data into different types.


Type specifies how the object is stored in memory.

Type answers the question "What is this thing?"

The `typeof()` function returns an object's type:
```{r}

```


## Classes

Every object also has at least one class.

Class answers the question "How does this thing behave?"

The `class()` function returns an object's class:
```{r}

```

Classes are more important than types for day-to-day programming.


## Identifying Vectors

Technically, class and type are independent.

For vectors, usually the class and type are the same:
```{r}

```

Vectors of decimal numbers are an exception:
```{r}

```

## Implicit Coercion

R can automatically convert or **coerce** types in one direction:

    logical -> integer -> numeric -> complex -> character


For example:
```{r}

```

This process is called **implicit coercion**.


We already saw this for the `c()` function:
```{r}

```

There are data types R will never implicitly coerce:
```{r}

```













Matrices, Arrays, & Lists
=========================

## Matrices & Arrays

**Matrices** are two-dimensional containers for values.

You can create a matrix from a vector with the `matrix()` function:
```{r}

```

The elements of a matrix must all have the same type.

It's usually faster to operate on matrix columns rather than rows.

The matrix multiplication operator is `%*%`:
```{r}

```


**Arrays** generalize vectors and matrices to higher dimensions.

Use the `array()` function to create an array:
```{r}

```


## Lists


A **list** is a container for elements with _different_ types.

The `c()` function returns a list if implicit coercion is not possible:
```{r}

```

You can also create a list with the `list()` function:
```{r}

```

List elements can have names:
```{r}

```

Vectorized functions don't work for lists:
```{r}

```

Access list elements by position with `[[`, the extraction operator:
```{r}

```

















Special Values
==============

R has four special values.


## Missing Values

`NA` represents a **missing value** in a data set:
```{r}

```

The missing value `NA` is contagious!
```{r}

```

Using a unknown argument in a computation usually produces an unknown result.


## Null

`NULL` represents a value that's not defined _in R_.

`NULL` usually indicates absence of a result:
```{r}

```

For instance, if we try to get the matrix dimensions of a vector:
```{r}

```


## Not a Number

`NaN`, or "not a number", represents a value that's not defined mathematically.


```{r}

```


# Infinite Values

`Inf` represents infinity. Produced by some computations:

```{r}

```





















Making Comparisons
==================

We saw operators for doing arithmetic: `+`, `-`, `*`, `/`, `^`

R also has operators for making comparisons:

* `==` for equality
* `<`, `<=`, `>`, `>=` for inequality

For example:
```{r}

```

Comparison operators are vectorized, just like arithmetic operators:
```{r}

```

To check equality for special values, use the `is.` functions:

* `is.na()`
* `is.null()`
* `is.nan()`
* `is.infinite()`

For example:
```{r}

```


## Equality & Numerical Precision

To check equality within a tolerance, use `all.equal()`:
```{r}

# isTRUE

```

Two reasons to use `all.equal()`:

1. Sometimes you want a tolerance (e.g., simulations).

2. On computers, all numbers are discrete (so most are imprecise).

As an example of the second point:
```{r}

```


To check whether two objects are identical, use `identical()`.

The `identical()` function is _not_ vectorized:
```{r}

```
