---
title: "Stat 33B - Lecture Notes 2"
date: August 30, 2020
output: pdf_document
---

Data Types
==========

In statistics, we categorize data into different types:

* Continuous (real numbers)
* Discrete (integers, or finite number of values)
* Logical (1 and 0s, T and Fs)
* Nominal (categorical values with no ordering)
* Ordinal (categorical values with ordering) (good, better, best)
* Graph (network data)
* Textual (books, websites, etc)

R also categorizes data into different types.


Type specifies how the object is stored in memory.

Type answers the question "What is this thing?"

The `typeof()` function returns an object's type:
```{r}
typeof(5) # double precision floating point

typeof(6.1)

typeof("hi")

typeof(1 + 3i)

typeof(TRUE)

typeof(FALSE)

typeof(5L)

typeof(mean) # type of function
```


## Classes

Every object also has at least one class.

Class answers the question "How does this thing behave?"

The `class()` function returns an object's class:
```{r}
class(6)

class(TRUE)

class(0 + 3i)

class(mean)
```

Classes are more important than types for day-to-day programming.


## Identifying Vectors

Technically, class and type are independent.

For vectors, usually the class and type are the same:
```{r}
class(TRUE)
typeof(TRUE)

class(c(0 + 1i, 0 + 5i))
typeof(0 + 1i)
```

Vectors of decimal numbers are an exception:
```{r}
class(5.1)
typeof(5.1)
```

## Implicit Coercion

R can automatically convert or **coerce** types in one direction:

    logical -> integer -> numeric -> complex -> character


For example:
```{r}
class(c(1, TRUE))

c(TRUE, 0 + 3i)
```

This process is called **implicit coercion**.


We already saw this for the `c()` function.

We can also use implict coercion with other functions:
```{r}
TRUE + 1
```

There are data types R will never implicitly coerce:
```{r}
c(1, sin)
```



Matrices, Arrays, & Lists
=========================

## Matrices & Arrays

**Matrices** are two-dimensional containers for values.

You can create a matrix from a vector with the `matrix()` function:
```{r}
matrix(c(1, 3, 7, 9), 2) # values filled in in column order

# data types of elements have to be same or will be coerced
m = matrix(c(TRUE, FALSE), 1)
m

class(m)

# type of matrix is same as type of elements
typeof(m)
```

The elements of a matrix must all have the same type.

It's usually faster to operate on matrix columns rather than rows.

The matrix multiplication operator is `%*%`:
```{r}
m1 = matrix(seq(1, 8), 2)
m1

m2 = matrix(seq(1, 12), 4)
m2

m1 %*% m2
```

Access matrix elements with `[`, the subset operator:
m[row, column]
```{r}
m1[2, 3]
m1

# get the whole first row as a vector
m1[1, ]

# get the whole column as a vector
m1[, 3]
```


**Arrays** generalize vectors and matrices to higher dimensions.

Use the `array()` function to create an array:
```{r}
x = array(seq(1, 8), c(2, 2, 2))
x

class(x)
typeof(x) # same as element type
```

Like matrices, access array elements with `[`, the subset operator.



## Lists


A **list** is a container for elements with _different_ types.

The `c()` function returns a list if implicit coercion is not possible:
```{r}
class(c(sin, 1))
typeof(c(sin, 1))
```

You can also create a list with the `list()` function:
```{r}
list(1, 2, 3)

list(1, sin, TRUE)
```

List elements can have names:
```{r}
list(x = 3)
```

Vectorized functions don't work for lists:
```{r}
list(1, 2) + list(3, 4)
```

Access list elements by position with `[[`, the extraction operator:
```{r}
mylist = list(1, 2, TRUE)
mylist

mylist[[3]]
```



Special Values
==============

R has four special values.


## Missing Values

`NA` represents a **missing value** in a data set:
```{r}
NA

class(NA)
typeof(NA)

x = c(1, NA, 2)
class(x)

x = c("hi", NA)
class(x)
x

c("NA", NA)
```

The missing value `NA` is contagious!
```{r}
NA + 3

3 * NA

7 / NA
```

Using a unknown argument in a computation usually produces an unknown result.


## Null

`NULL` represents a value that's not defined _in R_.

`NULL` usually indicates absence of a result:
```{r}
c()

dim(c(1, 2))

class(NULL)
typeof(NULL)

c(1, NULL)
```

For instance, if we try to get the matrix dimensions of a vector.


## Not a Number

`NaN`, or "not a number", represents a value that's not defined mathematically.


```{r}
NaN

class(NaN)
typeof(NaN)

c(1, NaN)

0 / 0
```


# Infinite Values

`Inf` represents infinity. Produced by some computations:

```{r}
13 / 0

Inf

c(1, Inf)

class(Inf)
typeof(Inf)
```



Making Comparisons
==================

We saw operators for doing arithmetic: `+`, `-`, `*`, `/`, `^`

R also has operators for making comparisons:

* `==` for equality
* `<`, `<=`, `>`, `>=` for inequality

For example:
```{r}
1 < 2

1 == 1

5 <= -1
```

Comparison operators are vectorized, just like arithmetic operators:
```{r}
c(1, 2, 3) == c(4, 5, 3)

c("hi", "hello") == "hi"

c(1, 2) < c(3, 4)

NA == 3

NA == NA
```

To check equality for special values, use the `is.` functions:

* `is.na()`
* `is.null()`
* `is.nan()`
* `is.infinite()`

For example:
```{r}
is.na(NA)

is.na(c(1, NA, 3))

is.null(NULL)

is.null(NA)
```


## Equality & Numerical Precision

To check equality within a tolerance, use `all.equal()`:
```{r}
5 == 5

all.equal(5, 5)

all.equal(7, 21)

# isTRUE
isTRUE(3)

isTRUE(all.equal(7, 21))
isTRUE(all.equal(7, 7))
```

Two reasons to use `all.equal()`:

1. Sometimes you want a tolerance (e.g., simulations).

2. On computers, all numbers are discrete (so most are imprecise).

As an example of the second point:
```{r}
0.1 == 0.3 / 3

0.3 / 3

all.equal(0.1, 0.3 / 3)
```


To check whether two objects are identical, use `identical()`.

The `identical()` function is _not_ vectorized:
```{r}
c(6, 7) == c(6, 7)

identical(c(6, 7), c(6, 7))

identical(c(6, 7), c(6L, 7L))
```
