---
title: "STAT 33B Workbook 1"
date: "Sep 3, 2020"
author: "Ming Fong (3035619833)"
output: pdf_document
---

This workbook is due __Sep 3, 2020__ by 11:59pm PT.

The workbook is organized into sections that correspond to the lecture videos
for the week. Watch a video, then do the corresponding exercises _before_
moving on to the next video.

Workbooks are graded for completeness, so as long as you make a clear effort to
solve each problem, you'll get full credit. That said, make sure you understand
the concepts here, because they're likely to reappear in homeworks, quizzes,
and later lectures.

As you work, write your answers in this notebook. Answer questions with
complete sentences, and put code in code chunks. You can make as many new code
chunks as you like.

In the notebook, you can run the line of code where the cursor is by pressing
`Ctrl` + `Enter` on Windows or `Cmd` + `Enter` on Mac OS X. You can run an
entire code chunk by clicking on the green arrow in the upper right corner of
the code chunk. 

Please do not delete the exercises already in this notebook, because it may
interfere with our grading tools.


Data Types
==========

Watch the "Data Types" lecture video.

## Exercise 1

In R, if you pass vectors with different lengths to a binary operator, the
shorter one will be __recycled__. This means the elements of the shorter vector
will be repeated to match the length of the longer vector.

Use the recycling rule to explain what's happening in each of these lines of
code:
```{r}
c(1, 2) + c(3, 4)

c(20, 30, 40) / 10

c(1, 3) + c(0, 0, 0, 0, 0)
```

YOUR ANSWER GOES HERE:

1. Two vectors of equal length are added. The value at the nth index of the first array is added to the nth index of the second array.
2. The value of each index is divided by 10.
3. The values of the first vector are copied until there are the same number of elements in each vector. Then the vectors are added like in the first example.

## Exercise 2

Run each line in the following code chunk and inspect the result. For each one,
state the type and class of the result, and explain why the result has that
type.

```{r}
c(TRUE, "hello", 3, 6)

3L + 3i

c(3L, 4L, 5L) / TRUE
```

YOUR ANSWER GOES HERE:

1. 
2. 
3. 

## Exercise 3

Another way to create vectors is with the `rep()` function. The `rep()`
function creates a vector by replicating a value or vector of values.

1. The first parameter of `rep()` is the thing to replicate. The second
   parameter, `times`, is the number of times to to replicate. Use `rep()` to
   make a vector with 10 elements, all equal to 78.

2. What happens if you pass a vector as the first argument to `rep()`? Give
   some examples.

3. Skim the help file `?rep`. What happens if you pass a vector as the second
   argument to `rep()`? The help file might seem a bit cryptic, so you'll also
   need to experiment. Give some examples.

YOUR ANSWER GOES HERE:



## Exercise 4

Yet another way to create vectors is with the `seq()` function. The `seq()`
function creates a vector that contains a sequence of numbers.

Skim the help file `?seq`. Give some examples of creating vectors with the
`seq()` function.

YOUR ANSWER GOES HERE:



## Exercise 5

In R, `T` and `F` are shortcuts for `TRUE` and `FALSE`.

1. What happens if you try to assign a value to `TRUE`?

2. What happens if you try to assign a value to `T`?

3. Check that what you observed in #1 and #2 is also true for `FALSE` and `F`.
   Why might it be safer to use `TRUE` and `FALSE` rather than `T` and `F` in
   code?

YOUR ANSWER GOES HERE:



Matrices, Arrays, & Lists
=========================

Watch the "Matrices, Arrays, & Lists" lecture video.

## Exercise 6

Recall that many of R's functions are vectorized, which means they are applied
element-by-element to vectors.

1. What happens if you call a vectorized function on a matrix?

2. What happens if you call a vectorized function on an array?

Give examples to support your answer.

YOUR ANSWER GOES HERE:



## Exercise 7

Suppose we want to multiply a length-2 vector with a 2-by-2 matrix.

What happens if you use `*` to multiply them? What happens if you use `%*%`?

Give some examples that show the difference, including for vectors and matrices
of other sizes.

YOUR ANSWER GOES HERE:



## Exercise 8

The `c()` function combines vectors, but it can also combine lists. Use
`list()` to create two lists, and show that `c()` can be used to combine them.

YOUR ANSWER GOES HERE:



Special Values
==============

Watch the "Special Values" lecture video.

## Exercise 9

Skim the help file for the `mean()` function.

1. What happens if you call the mean function on a vector that contains missing
   values? Is there a way to override this behavior?

2. What happens if you call the mean function on a vector that contains `NaN`
   values or infinite values?

In each case, provide examples to suport your answers.

YOUR ANSWER GOES HERE:



Making Comparisons
==================

Watch the "Making Comparisons" lecture video.

## Exercise 10


Each of the following lines of code produces a result that, at a glance, you
might not expect. Explain the reason for each result.
```{r}
3 == "3"

50 < '6'

isTRUE("TRUE")
```

YOUR ANSWER GOES HERE:



## Exercise 11

Suppose you want to check whether any of the values in `c(1, 2, 3)` appear in
the vector `c(4, 1, 3, 1)`.

Novice R users often expect they can check with the code:
```{r}
c(1, 2, 3) == c(4, 1, 3, 1)
```

1. Explain why the code above is not correct, and what's actually happening.

2. The correct way is to use the `%in%` operator. Give some examples of using
   the `%in%` operator. Recall that you can access its help page with
   `?"%in%"`.

YOUR ANSWER GOES HERE:



Submitting Your Work
====================

Congratulations, you made it through the first workbook!

You need to submit your work in two places:

* Submit this Rmd file with your edits on bCourses.
* Knit and submit the generated PDF file on Gradescope.
